---
title: "Analyzing pig movement across populations"
author: Mark Wilber and Sarah Chinn
output:
  html_notebook: default
  pdf_document: default
bibliography: /Users/mqwilber/Dropbox/Documents/Bibfiles/Projects_and_Permits-feral_swine.bib
---

## The questions of the analysis

1. How do pigs select for anthropogenic forage resources and natural forage resources on the landscape?
2. How does the selection for anthropogenic forage resources and natural forage resources vary temporally? Can this be described by climatic variables such as snowdepth and temperature? 
3. Do anthropogenic and natural forage resources interact to influence pig movement on a landscape? I.e. when both are present, does the availability of one affect the use of the other?

The goal of this analysis is **not** to build the best possible model of pig movement on a landscape.  Rather, it is to understand how forage resources influence a resource selection in pigs whether this varies in a predictable way between populations.

To answer these questions, we are going to use the framework of continuous-time, discrete-space movement movels described in the previous summary I sent around. The analysis below is performed on 5 populations. 

### Variable descriptions

The analyses below includes the following covariates.  A cell is 30m by 30m.

1. *Correlated random walk* (`crw`): A gradient-response that determines the propensity of a pig to move in the direction that is was moving on it previous move. A positive coefficient indicates that a pig tends to move in the same direction as their previous step.
2. *Canopy cover* (`canopycover_loc` and `canopycover_grad`): The tree density of a given cell.  `canopycover_loc` is a location-based covariates and a negative coeeficient indicates that as cell cover increases the rate at which a pig leaves that cell decreases (i.e. the time that the pig spends in that cell increases). `canopycover_grad` is a gradient-based covariate wand a positive coefficient indicates that pigs tend to move in the direction of a increasing canopy cover relative to their current position. 
3. *Distance to water* (`water_loc` and `water_grad`): The distance a pig is from a perennial water source.  `water_loc` is a location-based covariate where a positive coefficient indicates that pigs spend less time in cells that are further from water. `water_grad` is a gradient-based coefficient where a negative coefficient means that pigs tend to move closer to water.
4. *Distance to crops* (`crop_loc` and `crop_grad`): Distance to the neareast crop field. `crop_loc` is a location-based covariate where a positive coefficient indicates that pigs spend less time in cells that are further way from crop fields.  `crop_grad` is a gradient-based covariate where a negative coefficient means that pigs tend to move closer to crop fields.
5. *Masting tree density* (`masting_loc` and `masting_grad`): The density of masting trees in a cell. `masting_loc` is a location-based covariate where a negative coefficient indicates that a pig spends more time in a cell with higher masting tree density. `masting_grad` is a gradient-based covariate where a positive value indicates that pigs move in the direction of higher masting tree density, relative to where they currently are.
6. *NDVI* (`ndvi_loc` and `ndvi_grad`): A proxy for plant productivity/natural plant forage not a accounted for by masting trees. Varies on a monthly scale.  `ndvi_loc` is a location-based covariate where a negative coefficient indicates that pigs spend more time in cells with higher NDVI.  `ndvi_grad` is a gradient-based covariate where a positive coefficient means that pigs tend to move in the direction of increasing NDVI, relative to where they currently are.
7. *Temperature* (`temperature_loc`): Mean monthly temperature varying by month. Interacts with precipitation, cover, water, and forage covariates.
8. *Precipitation* (`precipitation_loc`): Total monthly precipitation.  Interacts with temperature, cover, water, and forage covariates.
9. *Snow depth* (`snowdepth_loc` and `snowdepth_grad`): Monthly variation in snowdepth at a 1km by 1km scale. `snowdepth_loc` is a location-based covariate where a positive coeffecient indicates that pigs spend less time in cells with higher snow depth.  `snowdepth_grad` is a gradient-based coefficient where a negative value indicates that pigs tend to move away from increasing snowdepth, relative to their current location.

Of these variables *NDVI*, *Temperature*, *Precipication*, and *Snow depth* vary monthly, *Distance to crops* can vary yearly, and *Masting tree density*, *Canopy cover*, and *distance to water* are fixed.

## Data summary

Here are some brief summary plots of the populations that are being analyzed. One plot looks at median location of pigs in space and the other gives the break down between male and female pigs in a study.

```{r}
library(data.table)
library(glmnet)
library(foreach)
library(ggplot2)
require(doMC)
```

```{r}
tejondat = fread("/Users/mqwilber/Repos/rsf_swine/results/glmdata_by_study/tejon.csv")
txdat = fread("/Users/mqwilber/Repos/rsf_swine/results/glmdata_by_study/txcamp.csv")
fldat = fread("/Users/mqwilber/Repos/rsf_swine/results/glmdata_by_study/fl_raoul.csv")
txtylerdat = fread("/Users/mqwilber/Repos/rsf_swine/results/glmdata_by_study/tx_tyler_w2.csv")
srelcdat = fread("/Users/mqwilber/Repos/rsf_swine/results/glmdata_by_study/srel_contact.csv")


# Add in sex data
pigattrib = fread("~/Repos/rsf_swine/data/formatted/pig_attributes.csv")[, list(pigID, sex)]

# Check NA values, should have some in the crop values
tejondat[, lapply(.SD, function(x) any(is.na(x)))][1,] == FALSE
txdat[, lapply(.SD, function(x) any(is.na(x)))][1, ] == FALSE

# Merge the pig attributes to the data
tejondat = merge(tejondat, pigattrib, by="pigID")[!is.na(ndvi_loc)]
txdat = merge(txdat, pigattrib, by="pigID")[!is.na(ndvi_loc)]
fldat = merge(fldat, pigattrib, by="pigID")[!is.na(ndvi_loc)]
txtylerdat = merge(txtylerdat, pigattrib, by="pigID")[!is.na(ndvi_loc)]
srelcdat = merge(srelcdat, pigattrib, by="pigID")[!is.na(ndvi_loc)]
```

```{r}
pig1 = tejondat[pigID == "tejonM302"]

res = build_design_matrix(pig1, c("crop"), modeltype="main_effects")
pig1 = res$data

evalform = res$evalform
evalform = update.formula(evalform, ~ . - sex)

Xmat = model.matrix(evalform, data=pig1)

registerDoMC(cores=4)

# Fit overall model with GLM LASSO
fit_model1 = cv.glmnet(Xmat, y=pig1$z, offset=log(pig1$tau), family="poisson", 
                       alpha=1, nlambda=20, nfolds=4, parallel=TRUE)


```

```{r}
plot(fit_model1)
fit_model1$glmnet.fit$beta[, 10, drop=F]
```


```{r}
source("pigfxns.R")
dataset = list(txcamp=txdat, tejon=tejondat, fl_raoul=fldat, tx_tyler_w2=txtylerdat, srel_contact=srelcdat)
```

```{r}
for(studynm in names(dataset)){

  dat = dataset[[studynm]]
  meandat = dat[, list(meanlat=median(y.adj), meanlon=median(x.adj), sex=unique(sex)[1]), by=pigID]
  plot1 = ggplot(meandat) + geom_point(aes(x=meanlon, y=meanlat, color=pigID)) + theme_bw() + xlab("Longitude") + ylab("Latitude") + ggtitle(studynm)
  plot2 = ggplot(meandat) + geom_bar(aes(x=sex)) + theme_bw() + ggtitle(studynm)
  print(plot1)
  print(plot2)

}
```

# Analyzing the influence of foraging resources on the movement patterns of pigs

A four step analysis. Each analysis is performed separately for each population.

1. Analyze a simple main effect model with no interactions
2. Analyze a GAM model where use of resources varies both daily and monthly
3. Analyze a main effect model where with monthly and daily variation in movement and interactions between foraging resources.
4. Analyze a model in which movement varies daily and seasonal variation in foraging is driven by monthly variation in temperature, precipitation, and snowdepth.

**Goal**: How do anthropogenic (i.e. crops) and natural forage resources affect pig movement, how do these effects vary over time, and how do these effects vary with climatic variables such as temperature and precipitation?

### Model 1: Simple main effects model

This model is giving us a sense of overall patterns in how pigs are using foraging resources, cover, and water. Purposefully ignoring a lot of the inherent complexity in this first set of models.

Fit the GLM LASSO model to each dataset

```{r}
require(doMC)

allmodel1_fits = list()
allXs = list()

for(studynm in names(dataset)){
  
  cat("Fitting", studynm, "\n") 
  dat = dataset[[studynm]]
  
  # Following Buderman et al. standardize on a per pig level, rather than across all pigs?
  res = build_design_matrix(dat, c("crop"), modeltype="main_effects")
  dat = res$data
  dataset[[studynm]] = res$data
  
  evalform = res$evalform
  
  Xmat = model.matrix(evalform, data=dat)
  allXs[[studynm]] = Xmat
  
  registerDoMC(cores=4)
  
  # Fit overall model with GLM LASSO
  fit_model1 = cv.glmnet(Xmat, y=dat$z, offset=log(dat$tau), family="poisson", 
                       alpha=1, nlambda=20, nfolds=4, parallel=TRUE)
  allmodel1_fits[[studynm]] = fit_model1
}
```

Examine the results of the regularized model fit. 

```{r}

mod1_betas = list()
for(studynm in names(allmodel1_fits)){
  tfit = allmodel1_fits[[studynm]]
  # plot(tfit)
  print(tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.min), drop=F])
  mod1_betas[[studynm]] = tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.1se), drop=T]
}
```

Plot results by coefficient.

```{r, fig.height=8, fig.width=10}

tmpdt = list()
for(studynm in names(mod1_betas)){
  
  betas = mod1_betas[[studynm]]
  nms = names(betas)
  dt = data.table(coef=nms, beta=betas, study=studynm)
  tmpdt[[studynm]] = dt
}

mod1betas = do.call(rbind, tmpdt)
ggplot(mod1betas) + geom_point(aes(x=study, y=beta, color=study), size=4) + facet_wrap(~coef,  scales='free') + 
              geom_hline(aes(yintercept=0)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size=8))

```


**Summary for Model 1**

*Consistent effect*

1. `canopycover_loc`: For all studies, pigs tended to spend more time in higher cover habitat relative to lower cover habitat
2. `canopycover_grad`: For all studies, either showed no directional bias or tended to move toward higher cover habitat.
3. `crw`:  All pigs tended to continue moving in the direction they were previously moving.

*Variable effects*


## Model 2:  Time-varying analysis for studies

**Question:** What is the effect of time of day and season on the resource selection of pigs? 

- Not including interactions between resources
- Not including monthly precipitation and tempurature variables as the montly time variation in resource use should (less mechanistically) account for this
- Explore daily and monthly changes in resource use.
  *Daily effects*: General movement
  *Monthly efects*: Canopy cover, forage resources, and distance to water.  Note NDVI is already varying by month, so I am not sure it makes a ton of sense to model a time-varying coefficient for this variable. Doing it anyway for now. 


```{r}
source("pigfxns.R")
df_hour = 6 # Number of basis vectors for the splines for hour

# Allow the monthly basis functions to vary between studies as they are of different lengths.
# If a study spans the course of a year, use a cyclic spline, otherwise use a cubic regression spline
cc_year = list(txcamp=list(FALSE, "cr", 6), tejon=list(TRUE, "cc", 6), fl_raoul=list(FALSE, "cr", 4), tx_tyler_w2=c(FALSE, "cr", 1), srel_contact=list(FALSE, "cr", 6))

allmodel2_fits = list()
allXgams = list()

for(studynm in names(dataset)){
  
  cat("Fitting", studynm, "\n")
  dat = dataset[[studynm]]
  
  # Extract hour of day and month of year for spline fitting
  dat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
  dat[ , hourofday:=hour(datetime)]
  dat[ , monthofyear:=month(datetime)]
  
  Xgam  = build_gam_design_matrix(allXs[[studynm]], dat, c("(Intercept)", "crw_z", "sexM"), df_hour=df_hour, df_month=cc_year[[studynm]][[3]], cyclic_year=cc_year[[studynm]][[1]])
  allXgams[[studynm]] = Xgam
  
  registerDoMC(cores=4)
  fitmodel2 = cv.glmnet(x=Xgam, y=dat$z, offset=log(dat$tau), 
                        family="poisson", alpha=1, nlambda=20, nfolds=4, parallel=TRUE)
  allmodel2_fits[[studynm]] = fitmodel2
}
```

Explore the nature of the time-varying effects after LASSO regularization.  Extract the betas that minimize CV Deviance. 

```{r}

gambetas = list()
for(studynm in names(allmodel2_fits)){
  
  tfit = allmodel2_fits[[studynm]]
  #plot(tfit)
  gambetas[[studynm]] = tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.min), drop=T]
}
```

After regularization, how do these effects vary with time?  Let's look at the daily effect on movement rate first


```{r}
library(ggplot2)

for(studynm in names(dataset)){
  
  dat = dataset[[studynm]]
  splinehour = s(hourofday, bs="cc", k=df_hour)
  pred_hours = Predict.matrix(smooth.construct2(splinehour, dat, NULL), data.frame(hourofday=0:23)) # B-spline basis for hour effect
  bestbetas = gambetas[[studynm]]
  
  dailyres = list()
  for(pattern in c("hour_*")){
    
    eff = pred_hours %*% t(t(bestbetas[names(bestbetas) %like% pattern]))
    effdt = data.frame(hour=0:23, beta=eff, coef=pattern)
    dailyres[[pattern]] = effdt
  }
  
  dailyres = do.call(rbind, dailyres)
  tplot = ggplot(dailyres) + geom_line(aes(x=hour, y=beta)) + xlab("Hour of day") + ylab("Daily Rate") + theme_bw() + facet_wrap(~coef) + ggtitle(studynm)
  print(tplot)
  
}
```


As expected, all pigs show reduced movement rates at some time of the day, but these effects are not consistent across pig populations.  A few populations tend to show reduced movement later in the day, while others show reduced movement around midday.

---

Now let's look at how our covariates vary with month.

```{r}

allmonthres = list()
for(studynm in names(dataset)){
  
  dat = dataset[[studynm]]
  
  maxdfmonth = ifelse(cc_year[[studynm]][[3]] == 1, 2, cc_year[[studynm]][[3]])
  splinemonth = s(monthofyear, bs=cc_year[[studynm]][[2]], k=ifelse(cc_year[[studynm]][[1]], maxdfmonth, maxdfmonth  - 1))
  
  minmonth = min(dat$monthofyear)
  maxmonth = max(dat$monthofyear)
  
  if(cc_year[[studynm]][[3]] > 1)
    pred_months = Predict.matrix(smooth.construct2(splinemonth, dat, NULL), data.frame(monthofyear=seq(minmonth, maxmonth, len=20)))
  else
    pred_months = t(t(rep(1, 20)))
  bestbetas = gambetas[[studynm]]
  
  monthres = list()
  for(pattern in c("cover_loc_*", "cover_grad_*", "crop_grad_*", "crop_loc_*", "water_loc_*", "water_grad_*",
                 "month_*", "masting_loc_*", "masting_grad_*", "ndvi_loc_*", "ndvi_grad_*")){
    
    eff = pred_months %*% t(t(bestbetas[names(bestbetas) %like% pattern]))
    effdt = data.frame(month=seq(minmonth, maxmonth, len=20), beta=eff, coef=pattern)
    monthres[[pattern]] = effdt
  }
  
  monthres = do.call(rbind, monthres)
  monthres$study = studynm
  allmonthres[[studynm]] = monthres
  tplot = ggplot(monthres) + geom_line(aes(x=month, y=beta)) + xlab("Month of year") + 
                  ylab("Effect on movement rate") + theme_bw() + facet_wrap(~coef, scales="free") + ggtitle(studynm)
  print(tplot)
  
}

allmonthres = do.call(rbind, allmonthres)
tplot = ggplot(allmonthres) + geom_line(aes(x=month, y=beta, color=study)) + xlab("Month of year") + 
                  ylab("Effect on movement rate") + theme_bw() + facet_wrap(~coef, scales="free") + ggtitle(studynm)
print(tplot)

```


Some interesting population-specific effects

**Tejon**

1. The Tejon population shows strong seasonal variation in selection for resources
2. During the summer months there is a tendency to move toward crops and move faster once near crops.
3. This effect diminshes in the winter months in which there is a tendency to spend more time in cells with increased density of masting trees.  In contrast, there is a tendency to spend less time in cells with high masting density in the summer.  


**Camp Bullis**

1. Relatively weak seasonal effects of resource selection

**Florida, `fl_raoul`**




## Model 3: Interactive effects of resources on pig movement

This model is examining 

In others words, given a pig is in or near a natural or agricultural resource, how does the presence of another resource (or distance to that resource) affect the time the pigs spends in the current resource.

For this analysis, we are going to keep the strong effects of hour and month on general movement patterns, but only consider interactions between static forage resources (i.e. the effects of resources are not time varying). 


```{r}
source("pigfxns.R")

allmodel3_fits = list()
allXints = list()

for(studynm in names(dataset)){
  
  cat("Fitting", studynm, "\n")
  
  dat = dataset[[studynm]]
  inter = build_design_matrix(dat, c("crop"), modeltype = "interactions")
  dataset[[studynm]] = inter$data
  evalform = inter$evalform
  
  Xgam = allXgams[[studynm]]
  
  maxdfmonth = ifelse(cc_year[[studynm]][[3]] == 1, 2, cc_year[[studynm]][[3]])
  Xint = cbind(model.matrix(evalform, inter$data), Xgam[, c(paste0("hour_", 1:(df_hour - 1)), paste0("month_", 1:(maxdfmonth - 1)))])
  allXints[[studynm]] = Xint
  
  fitmodel3 = cv.glmnet(Xint, y=dat$z, offset=log(dat$tau), 
                    family="poisson", alpha=1, nlambda=20, nfolds=4, parallel=TRUE)
  allmodel3_fits[[studynm]] = fitmodel3
  
}
```

```{r}

mod3betas = list()
for(studynm in names(allmodel3_fits)){
  
  tfit = allmodel3_fits[[studynm]]
  #plot(tfit)
  print(studynm)
  print(tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.1se), drop=F])
  mod3betas[[studynm]] = tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.min), drop=T]
}
```

```{r, fig.width=10, fig.height=10}
tmpdt = list()

nm = 15
for(studynm in names(mod3betas)){
  
  betas = mod3betas[[studynm]][1:15]
  nms = names(betas)
  dt = data.table(coef=nms, betas, study=studynm)
  tmpdt[[studynm]] = dt
}

mod3betas_new = do.call(rbind, tmpdt)
ggplot(mod3betas_new) + geom_point(aes(x=study, y=betas, color=study), size=4) + facet_wrap(~coef,  scales='free') + 
              geom_hline(aes(yintercept=0)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size=8))
```

## Step 4: Replacing time-varying parameters with temperature, precipitation, and snow cover interactions

This will be similar to model 2, but instead of basis functions that vary by month (we will keep the daily basis function), we will replace these effect by interactions with temperature and precipitation. 

```{r}
source("pigfxns.R")

allmodel4_fits = list()
allXints_full = list()

for(studynm in names(dataset)){
  
  cat("Fitting", studynm, "\n")
  
  dat = dataset[[studynm]]
  inter = build_design_matrix(dat, c("crop"), modeltype = "full")
  dataset[[studynm]] = inter$data
  evalform = inter$evalform
  
  Xgam = allXgams[[studynm]]
  Xint = cbind(model.matrix(evalform, inter$data), Xgam[, paste0("hour_", 1:(df_hour -1))])
  allXints_full[[studynm]] = Xint
  
  fitmodel4 = cv.glmnet(Xint, y=dat$z, offset=log(dat$tau), 
                    family="poisson", alpha=1, nlambda=20, nfolds=4, parallel=TRUE)
  allmodel4_fits[[studynm]] = fitmodel4
  
}
```

```{r}
mod4betas = list()

for(studynm in names(allmodel4_fits)){
  
  tfit = allmodel4_fits[[studynm]]
  #plot(tfit)
  print(studynm)
  print(tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.1se), drop=F])
  mod4betas[[studynm]] = tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.1se), drop=T]
}
```

```{r, fig.height=10, fig.width=15}
tmpdt = list()

for(studynm in names(mod3betas)){
  
  betas = mod4betas[[studynm]][1:36]
  nms = names(betas)
  dt = data.table(coef=nms, betas, study=studynm)
  tmpdt[[studynm]] = dt
}

mod4betas_new = do.call(rbind, tmpdt)
ggplot(mod4betas_new) + geom_point(aes(x=study, y=betas, color=study), size=4) + facet_wrap(~coef,  scales='free') + 
              geom_hline(aes(yintercept=0)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, size=8))
```


Plot the temperature, precipitation, and NDVI effects over "month of year"

```{r}

for(studynm in names(dataset)){
  
  dat = dataset[[studynm]]
  meanvals = dat[, list(meantemp=mean(temperature_loc_z), meanprecip=mean(precipitation_loc_z), meanndvi=mean(ndvi_loc_z)), by=monthofyear]
  tplot = ggplot(meanvals) + geom_line(aes(x=monthofyear, y=meanndvi, color="NDVI")) + geom_line(aes(x=monthofyear, y=meantemp, color="Temperature")) + 
                        geom_line(aes(x=monthofyear, y=meanprecip, color="Precipitation")) + theme_bw() + ylab("Standardized value") + xlab("Month of year") + 
                        ggtitle(studynm)
  print(tplot)

}

```


There is less spatial variability in "natural" forage resources in Camp Bullis than in Tejon, which could diminish the effect on movement. Could be an interesting population-level predictor to consider for why the the role of natural vs crop resources on pig movement varies between population. The hypothesis is that with more homogeneous resource distributions (at what scale?), there is less signature of resource selection as pigs can wander more freely without worrying about getting to or staying in a particular resource patch. In general, the effects of all resources seem to be less in Camp Bullis, and I would suspect that has to do with the generally availability of resource (water and forage), though it would be nice to confirm this.  If each chosen cell is "as good as" the next, then resource selection should have little effect on pig movement and movement should be dominated by a random walk or other biological factors (escape from predation?).  This can also be a function of scale, where increasing spatial resolution can increase homogeneity and decrease the signature of foraging resource selection. 


---

## Post-hoc population-level analyses

Ok, so we have estimated the effects of various foraging resources on pig movement among different populations, now the goal is to ask how there 

```{r}

sumdats = list()
for(studynm in names(dataset)){
  
  dat = dataset[[studynm]]
  sumdat = dat[z == 1, lapply(list(ndvi=ndvi_loc, crop=crop_loc, masting=masting_loc, water=water_loc, dev=developed_loc), function(x) quantile(x, c(0.5)))]
  sumdat$study = studynm
  sumdats[[studynm]] = sumdat
  
}

sumdatsdt = do.call(rbind, sumdats)

```


```{r}
sindx = function(x, h){
  Im(sin(x)) / h
}

sindx(complex(real=2, imaginary = 10^-50), h=10^-50)
-cos(2)

```