---
title: "R Notebook"
output: html_notebook
---

Testing the CTCRW model on a pig trajectory using the crawl package

## Simulating a CTCRW model using the framework described in Johnson et al. 2008, Ecology

The model is formulated as a state space Markov model.

Define the parameters and matrices of the model

```{r}

get_T = function(beta, delta){
  # Get the T matrix of the state space representation
  
  Tmat = matrix(c(1, (1 - exp(-beta*delta)) / beta, 
                 0,   exp(-beta*delta)), nrow=2, ncol=2, 
                byrow=TRUE)
  return(Tmat)
}

get_Q = function(beta, sigma, delta){
  # Get the Q matrix of the state-space formulation
  
    # beta : temporal auto-correlation parameter. Bigger beta, less directional persistence
  # delta : time step
  # sigma : variability in velocity
  
  # Velocity variance
  zeta_var = sigma^2*(1 - exp(-2*beta*delta)) / 2*beta
  
  # Location variance
  xi_var = (sigma^2 / beta^2)*(delta - (2/beta)*(1 - exp(-beta*delta)) +
                                 (1 / (2*beta))*(1 - exp(-2*beta*delta)))
  
  # Covariance between velocity and location
  vl_covar = (sigma^2 / (2*beta^2))*(1 - 2*exp(-beta*delta) +
                                    exp(-2*beta*delta))
  
  Q = matrix(c(xi_var, vl_covar, vl_covar, zeta_var), nrow=2, ncol=2,
             byrow = TRUE)
  return(Q)
  
}

```

Simulate the random walk

```{r}

# First two are location and velocity in one direction, second two are
# location and velocity in another direction. 2D vector
start_locvel = c(20, 0, 20, 0)
beta = 15
sigma_proc = 3
sigma_obs = 0
delta = 1 # Time step
steps = 1000

true_locs = array(NA, dim=c(4, steps + 1))
true_locs[, 1] = start_locvel
obs_locs = array(NA, dim=c(2, steps + 1))

# These are for one dimension
Tmat = get_T(beta, delta)
Qmat = get_Q(beta, sigma_proc, delta) # location and velocity vcov
eta_mean = rep(0, 4) # location and velocity process error

# Observation matrix
zeromat = matrix(0, nrow=2, ncol=2)
Zvect = array(c(1, 0), dim=c(2, 1))

# Build the 4 dimensional vectors and matrices...why?  Because we are moving independently in two dimenions with location
# and velocity.
Tfull = rbind(cbind(Tmat, zeromat), cbind(zeromat, Tmat)) 
Qfull = rbind(cbind(Qmat, zeromat), cbind(zeromat, Qmat))
Zfull = rbind(c(Zvect, 0, 0), c(0, 0, Zvect))

for(i in 2:(steps + 1)){
  
  alpha_next = Tfull %*% true_locs[, i - 1] + MASS::mvrnorm(1, eta_mean, Qfull)
  true_locs[, i] = alpha_next 
  obs_locs[, i] = Zfull %*% alpha_next + t(t(MASS::mvrnorm(1, c(0, 0), diag(rep(sigma_obs^2, 2))))) 

}

```

```{r}
# Plot locations
locs = true_locs[c(1, 3), ]
plot(locs[1, ], locs[2, ])
lines(locs[1, ], locs[2, ])

points(obs_locs[1, ], obs_locs[2, ], col="red")
```

```{r}
# Format the data for fitting with crawl
sim_dat = data.frame(t(obs_locs[, 2:ncol(obs_locs)]))
sim_dat$time = 1:(steps)
colnames(sim_dat) = c("x", "y", "time")
sim_dat
```


```{r}
library(crawl)

# a is the initial value for the location vector
# P is the initial correlation matrix
initial = list(a=c(20, 0, 20, 0), 
               P=diag(c(10000^2,5400^2,10000^2,5400^2)))
fit_sim = crwMLE(mov.mod=~1, err.model=list(x=~1), data=sim_dat, Time.name="time", coord=c('x', 'y'), initial.state = initial, drift=FALSE)

```

```{r}
fit_sim
```

Try fitting the model with STAN


```{r}
library(rstan)
tl = t(true_locs)
deltas = rep(1, nrow(tl) - 1)

standata = list(N=nrow(tl), 
                obs_vals=tl,
                deltas=deltas)
saveRDS(standata, "temp_stan.rds")
stanfit = stan("stan_files/ctcrw_nobserr.stan", chains=3, data=standata, warmup=1000, iter=2000, cores=3)
```

```{r}
library(rethinking)
precis(stanfit)
```

Works great.  What if we have perfect observations of the locations but didn't observe the velocities? 

```{r}

just_loc = t(true_locs[c(1, 3), ])
standata2 = list(N = nrow(just_loc),
                 obs_vals=just_loc,
                 deltas=rep(1, nrow(just_loc) - 1))

fitstan2 = stan("stan_files/ctcrw_unobsvel.stan", data=standata2, warmup = 3000, iter=10000, cores=3, chains=3)

```

```{r}
precis(fitstan2)
traceplot(fitstan2, "beta")
```

This works in the sense that we get back the correct beta and sigma.  But man, it is a bit slow. 

Ok, let's scale this up to the last challenge (which is just not going to work well if the previous version struggled.).  We have imperfect locations and unobserved velocities.

```{r}

sobs = t(obs_locs)
sobs[1, ] = c(20, 20)
standata3 = list(N=nrow(sobs),
                 obs_vals = sobs,
                 deltas = rep(1, nrow(sobs) - 1))
fitstan3 = stan("stan_files/ctcrw.stan", data=standata3, chains=3, cores=3, warmup=2000, iter=5000)

```

```{r}
precis(fitstan3)
```

LOL!  AWFUL mixing.  And the traceplots show this too.  Interestingly, the means are pretty good.  We actually were able to recover. No way this is feasible for large datasets. 

```{r,message=FALSE}
predTime = seq(1, 21, len=100)

library(crawl)
predObj <- crawl::crwPredict(object.crwFit=fit_sim, 
                      predTime, 
                      speedEst=TRUE, 
                      flat=TRUE)
```


Now, view the predicted movement path

```{r,message=FALSE}
crwPredictPlot(predObj, "map")
```


## Empirical data analysis

**Goal**:  Fit the crawl model to some actual pig data.  Figure out how pigs move differently during the day. 

**0. Load in the data**

```{r}
library(data.table)
dat = fread("../data/formatted/pig_trajectories.csv")
dat$datetime = as.POSIXct(strptime(dat$datetime, format="%Y-%m-%d %H:%M:%S", tz="GMT"))
```

```{r}
library(ggplot2)
# Extract a single pigs trajectory
pig1 = dat[pigID == "20145txcamp"][1:1000]
ggplot(data=pig1) + geom_path(aes(x=longitude, y=latitude))
```

Look at distance a pig traveled

```{r}
library(geosphere)

get_distance_vect = function(longlat){
  # Calculate tghe distance that a pig traveled
  
  longs = longlat$longitude
  lats = longlat$latitude
  
  # Match distance vectors
  v1 = cbind(longs[-length(longs)], lats[-length(lats)])
  v2 = cbind(longs[-1], lats[-1])
  
  tuples = lapply(1:nrow(v1), function(i) list(v1[i, ], v2[i, ]))
  distvect = lapply(tuples, function(x) distm(x=x[[1]], y=x[[2]], fun=distHaversine))
  # 
  return(distvect)
}

pigdists = get_distance_vect(pig1[, c("longitude", "latitude")])
pigtimes = diff(pig1$datetime)
plot(cumsum(as.numeric(pigtimes)) / (60 * 24), cumsum(pigdists), type="l", xlab="Days")
```


```{r}
plot(pig1$datetime[1:(nrow(pig1) - 1)], pigdists, type="l")
```

Individual pigs show both seasonal movement variation and diurnal patterns.  Seems to be consistently reduced activity
during certain times of the day.  This is going to cause problems for the CTCRW model and we might need to induce some state shifting.


2. Project lat and long to the appropriate NAD measure

```{r}
library(sp)
library(rgdal)
coordinates(pig1) <- ~longitude+latitude
proj4string(pig1) <- CRS("+proj=longlat")

# This is projecting lat long to NAD83 in the appropriate rectangle that is centered on Michigan
pig1 <- spTransform(pig1, CRS("+init=epsg:3081"))
```

Plot the projection to see what the data actuall look like

```{r}
plot(pig1@coords[, 2], pig1@coords[, 1])
lines(pig1@coords[, 2], pig1@coords[, 1])
#print(nrow(short_dat))
```

3. Set the initial conditions

```{r}
library(crawl)
# Set the initial conditions for the location and movement vector. Location is at the starting location. Velocities start
# at 0.  P is the initial correlation matrix
initial = list(a=c(coordinates(pig1)[1,1],0,
                   coordinates(pig1)[1,2],0),
               P=diag(c(10000^2,5400^2,10000^2,5400^2)))

# Prior
ln.prior = function(theta){-abs(theta[4]-3)/0.5}
```

```{r}
displayPar(mov.model=~1,
           data=pig1)
```

```{r}
set.seed(123)
fit1 <- crwMLE(mov.model=~1,
               data=pig1, 
               Time.name="datetime",
               initial.state=initial,
               control=list(maxit=30, trace=0,REPORT=1),
               initialSANN=list(maxit=200, trace=0, REPORT=1))
```

```{r}
fit1
```


```{r,message=FALSE}
predTime = seq(min(pig1$datetime), max(pig1$datetime), len=100)

predObj <- crawl::crwPredict(object.crwFit=fit1, 
                      predTime, 
                      speedEst=TRUE, 
                      flat=TRUE)
```


Now, view the predicted movement path

```{r,message=FALSE}
crwPredictPlot(predObj, "map")
```


```{r}
simobj = crwSimulator(fit1)
samp = crwPostIS(simobj, fullPost=F)
```