---
title: "Exploratory data analysis for pig data"
output: html_notebook
---

# Basic data cleaning

```{r}
# Load in the pig data
library(data.table)
dt = fread("../data/formatted/full_pig_data.csv")
```

Check for missing lat-long values.  Hopefully the data cleaning has dealt with these...

```{r}
ind = is.na(dt$latitude) | is.na(dt$longitude)
sum(ind)
```

Recovert to datetime 

```{r}
dt$datetime = as.POSIXct(dt$datetime, tz="GMT")
```

## Some exploratory data analysis

```{r}
length(unique(dt$pigID))
```

```{r}
dt[, list(numpigs=length(unique(pigID))), by="study"]
```

We have 480 pigs, the `movepig` study site has 219 of these pigs.

Look at a plot of pigs in the USA

```{r}
library(ggmap)

meandat = dt[, list(meanlat = mean(latitude), meanlong = mean(longitude)), by="study"]
# All the pig data
us <- c(left = -125, bottom = 23.75, right = -67, top = 54)
map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
plt = ggmap(map) + geom_point(data=dt[pigID != "canada679059BF21"], aes(x=longitude, y=latitude, color=study), size=0.5)
plt
ggsave("../results/datamap.jpg", width=8, height=5)
```

Something is up with Kilgo! We will need to figure that out.  The Canada data also looks a little suspect...this is because the collars are recording when they are not on the animal.

```{r}
plt_can = ggmap(map) + geom_point(data=dt[dt$study == "canada" & dt$pigID == "canada679059B", ], aes(x=longitude, y=latitude, color=pigID), size=0.5)
plt_can
ggsave("../results/679059B.jpg", width=8, height=5)
```


```{r}
ggmap(map) + geom_point(data=dt[study == "tejon", ], aes(x=longitude, y=latitude))
```

Get the maximum time length at which fixings were taken for each individual pig

```{r}
fixinfo = dt[, list(study=unique(study),
                         deltat = max(datetime) - min(datetime), 
                         numfixes=length(datetime),
                         start=min(datetime),
                         end=max(datetime)), by=list(pigID)]
ggplot(data=fixinfo) + geom_histogram(aes(x=deltat))
```

```{r}
fixinfo[which.max(deltat), ]
```

So the Kilgo data is genearlly just wonky. Ignore it for now

```{r}
ggplot(data=fixinfo[study != "srs_kilgo", ]) + geom_histogram(aes(x=deltat))
```


Order by fix times.  

```{r}
fixinfo[order(deltat, decreasing = T)]
```

Large variation in how long pigs were collared for. 

```{r}
fixinfo[, fixrate:= numfixes / as.numeric(deltat)]
fixinfo[, timeperfix:= (as.numeric(deltat) / numfixes)*24*60]
fixinfo[order(timeperfix, decreasing = F), ]
ggplot(data=fixinfo[study != "srs_kilgo"]) + geom_histogram(aes(x=timeperfix))
```

---

Look at how collaring times vary across studies

```{r}
studies = unique(fixinfo$study)

plots = list()
for(i in 1:length(studies)){
  
  plots[[i]] = ggplot(data=fixinfo[study==studies[i]], aes(pigID)) + geom_linerange(aes(ymin=start, ymax=end)) + 
    facet_wrap(~study, scales = "free") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size=8))
}
```

```{r}
for(i in 1:length(plots)){
  print(plots[[i]])
}
```

There is some overlap in the times that pig were tagged within sutdies.  Not yet sure between studies though. Ideally, if our question is how pig movement changes across studies

Some summary stats by study

```{r}
# Order the GPS colors by pig and time 
dt = dt[order(pigID, datetime)]
diffs = dt[, list(deltat=mean(diff(datetime))), by=list(pigID)]

ggplot(diffs, aes(x=deltat / (60 * 24))) + geom_histogram() + xlab("Days")
```

Some pigs have a huge about of time between their GPS readings...

---

## Determining which sequences will be useful for movement analysis

Which pigs have a sequences of at least 10 where each fix is less than 1hr different
from the previous fix?

```{r}
source("pigfxns.R")

long_pigs = dt_trun[, list(longdat=runs(datetime), study=unique(study)), by=pigID]
print(sum(long_pigs$longdat))
print(nrow(long_pigs))
table(long_pigs$study[long_pigs$longdat])
```

212 pigs have at least 10 fixes in a row that are within 1 hr of each other. 

Let's just focus on these pigs for now. 

```{r}
# Subset data to only include pigs that have the right length of data
keys = long_pigs$pigID[long_pigs$longdat]

# Should be a cleaner and faster way to do this...perhaps with match
inds = rep(FALSE, nrow(dt_trun))
for(key in keys){
  inds = inds | (dt_trun$pigID == key)
}

# Remove the pigs that 
movedt = dt_trun[inds]
```


```{r}
table(movedt$study)
```


## Plot pig movement trajectories for each study

```{r}
library(ggmap)

# All the pig data
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(map) + geom_point(data=dt_trun, aes(x=longitude, y=latitude, color=study))
```

```{r}
# Pig data that will probably be useful for movement models
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(map) + geom_point(data=movedt, aes(x=longitude, y=latitude))
```


Look at a plot of pig movement in Texas.

```{r}
library(dplyr)
txcamp = movedt[pigID == "20125txcamp"]
map = get_googlemap(center=c(lon=-98.53675, lat=29.74411), zoom=14, maptype="hybrid") %>% ggmap()
map = map + geom_path(data=txcamp, aes(x=longitude, y=latitude), size=0.5)
map
```

```{r}
# Single pig diff
pigdiff = diff(txcamp$datetime)

inds = pigdiff < 100 # String of booleans
  
  # Count runs
run_vect = rle(inds)
run_vect
```

```{r}
write.csv(movedt, "../data/formatted/pig_trajectories.csv")
```
