---
title: "CTMC model for pigs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
# Load in necessary packages
library(data.table)
library(raster)
library(ggplot2)
library(lubridate)
library(crawl)
library(sp)
library(rgdal)
library(rgeos)
library(ctmcmove)
library(splines)

## Helper functions
source('/Users/mqwilber/Repos/rsf_swine/code/pigfxns.R')

```

For a single pig in the Camp Bullis data set, implement the movement method proposed by Hanks et al. 2015 to understand the resource utilization function.

## Step 1: Format the data and extract a single pig

```{r}

dt = fread("/Users/mqwilber/Repos/rsf_swine/data/formatted/full_pig_data.csv")

# Extract texas pigs
txpigs = dt[study == "txcamp"]
rm(dt)

# Convert datetime
txpigs$datetimef = as.POSIXct(strptime(txpigs$datetime, format='%Y-%m-%d %H:%M:%S'), tz="GMT")

# Extract pigs with the most observations
cntpigs = txpigs[, list(counts=length(datetime)), by=list(pigID)][order(counts, decreasing = T)]
head(cntpigs)


# Look for overlap
runsdt = txpigs[, list(runs=runs(datetimef, clength=500, ctime=30, ctimemin=0)), by="pigID"]
runsIDs = runsdt$pigID[runsdt$runs]

# extract the specific IDs
ind = array(FALSE, dim=nrow(txpigs))
for(ID in runsIDs){
	ind = ind | txpigs$pigID == ID
}

pigdt = txpigs[ind]
dim(pigdt)
unique(pigdt$pigID)

# Look for where the pigs are overlapping in time
minmax = pigdt[, list(min=min(datetimef), max=max(datetimef)), by="pigID"]
minmax

# Let's just focus on 03-03 for the following analysis
redpigs = minmax[month(min) == 3 & day(min) == 3]$pigID
redpigs

ind = array(FALSE, dim=nrow(pigdt))
for(ID in redpigs){
	ind = ind | pigdt$pigID == ID
}

# Get the test pigs
testpigsdt = pigdt[ind]
dim(testpigsdt)
pigIDs = unique(testpigsdt$pigID)
sum(table(testpigsdt$pigID))
```

```{r}
ind = (testpigsdt$pigID == pigIDs[3]) #(testpigsdt$pigID == testpigsdt$pigID[1])
pig1 = testpigsdt[ind][1:500, ]
pig1 = pig1[order(datetimef)]

```


## Step 2. Load in raster data

1. CropScape data for the area of interest

```{r}
plot(pig1$longitude, pig1$latitude)

# Get the bounding box for the pig data so we can apply it to the raster file
ymin = min(pig1$latitude)
ymax = max(pig1$latitude)
xmin = min(pig1$longitude)
xmax = max(pig1$longitude)

long1 = ggplot(data=pig1) + geom_line(aes(x=datetimef, y=longitude)) + theme_bw() + xlab("Time") + ylab("Longitude")
ggsave("../results/longplot.jpg", long1, width=5, height=3)
lat1 = ggplot(data=pig1) + geom_line(aes(x=datetimef, y=latitude)) + theme_bw() + xlab("Time") + ylab("Latitude")
ggsave("../results/latplot.jpg", lat1, width=5, height=3)
space = ggplot(data=pig1) + geom_path(aes(x=longitude, y=latitude)) + geom_point(aes(x=longitude, y=latitude)) + theme_bw() + xlab("Longitude") + ylab("Latitude")
ggsave("../results/latlong.jpg", space, width=5, height=3)

```

Load in a clip the cropscape raster file

```{r}
rcval = raster("../data/covariate_data/CDL_2016_clip_20180118163303_1712778437/CDL_2016_clip_20180118163303_1712778437.tif")

el = raster("~/Downloads/n30w099/imgn30w099_1.img")
# Crop the raster with some arbitrary buffer
buffer = 0.001
rcval1 = crop(rcval, extent(xmin - buffer, xmax + buffer, ymin - buffer, ymax + buffer))

# Project elevation raster 
elproj = projectRaster(el, rcval1)
elcrop = crop(elproj, extent(xmin - buffer, xmax + buffer, ymin - buffer, ymax + buffer))
rcval1
elcrop

jpeg("../results/landcover.jpg")
plot(rcval1)
points(pig1$longitude, pig1$latitude, cex=0.5)
dev.off()

jpeg("../results/elevation.jpg", width=500, height=500)
plot(elcrop)
points(pig1$longitude, pig1$latitude, cex=0.5)
dev.off()
```

```{r}
a = area(rcval1)
sqrt(values(a)[1] * 1000^2)
```

Close to 30 meter resolution on the grid cells.

Plenty of other rasters to consider, but let's start here.  Also, will need to convert these raster values to something more quantitative. Either zero or ones

```{r}
table(values(rcval1))
developed = forest = crop = water = dummy = rcval1
values(developed) = (values(developed) == 121) | 
                    (values(developed) == 122) | 
                    (values(developed) == 123) | 
                    (values(developed) == 124) | 
                    (values(developed) == 131)
        

values(forest)[values(rcval1) == 142] = 1 # Forest habitat
values(forest)[values(rcval1) != 142] = 0 # Not forest

values(water) = values(water) == 111
values(crop) = values(crop) < 100
values(dummy) = 0

```


```{r}
nms = c("developed", "forest", "crop", "water")
loc.stack = stack(forest, dummy) #stack(developed, forest, crop, water)
names(loc.stack) = c("forest_loc", "dummy_loc") #paste0(nms, "_loc")
grad.stack = stack(forest, dummy, elcrop) #stack(developed, forest, crop, water)
names(grad.stack) = c("forest_grad", "dummy_grad", "elev_grad") #paste0(nms, "_grad")
```

## Step 3: Fit a spline-based/interpolation-based continuous time movement model.

```{r}

# Scale lat and long for fitting
pig1$latsc = scale(pig1$latitude)
pig1$longsc = scale(pig1$longitude)

# Fitting an arbitrarily spline model just for now
# fitlong = lm(longsc ~ bs(datetimef, df=100), data=pig1)
# fitlat = lm(latsc ~ bs(datetimef, df=100), data=pig1)

fitlong = smooth.spline(as.numeric(pig1$datetimef), pig1$longsc) #lm(longsc ~ bs(datetimef, df=100), data=pig1)
fitlat = smooth.spline(as.numeric(pig1$datetimef), pig1$latsc) # lm(latsc ~ bs(datetimef, df=100), data=pig1)

iterpLong = approxfun(pig1$datetimef, pig1$longsc)
iterpLat = approxfun(pig1$datetimef, pig1$latsc)

ggplot(data=NULL) + geom_path(aes(x=predict(fitlong)$y, y=predict(fitlat)$y)) + geom_point(aes(x=pig1$longsc, y=pig1$latsc), color="red", size=0.5) + theme_bw() + 
                xlab("Scaled long.") + ylab("Scaled lat.")
ggsave("../results/movetraj.jpg", width=5, height=3)

plot(pig1$datetimef, pig1$longsc, type="l")
lines(pig1$datetimef, predict(fitlong)$y, col="red")

plot(pig1$datetimef, pig1$latsc, type="l")
lines(pig1$datetimef, predict(fitlat)$y, col="red")

```

Use the continous spline representations to build a continuous trajectory spaced at hour intervals over the entire course of the study

```{r}
mintime = min(pig1$datetimef)
maxtime = max(pig1$datetimef)
predTime = seq(mintime, maxtime, by="15 mins")
predlong =iterpLong(predTime)
predlat = iterpLat(predTime)

# For B-spline fit
# predlong =  predict(fitlong, newdata=data.frame(datetimef=predTime))
# predlat = predict(fitlat, newdata=data.frame(datetimef=predTime))

# predlong =  predict(fitlong, as.numeric(predTime))$y
# predlat = predict(fitlat, as.numeric(predTime))$y


# Convert back to the natural scale
predlongT = predlong*sd(pig1$longitude) + mean(pig1$longitude)
predlatT = predlat*sd(pig1$latitude) + mean(pig1$latitude)

# Make sure the projected paths don't go off the landscape

jpeg("../results/predmove.jpg")
plot(rcval1)
points(predlongT, predlatT, cex=0.1)
dev.off()

```

## Step 4: Set up the data for CTMC analysis

Following the steps in the Wilson et al., Calculating where each point is.
Pretty slow over all

```{r}

# Specify where the animal can move on the lanscape
move_field = rcval1
values(move_field) = 1

# Transition matrix
trans = transition(move_field, prod, 4)

# Convert the path to resident times...pretty slow
path = list(xy=cbind(predlongT, predlatT), t=as.vector(predTime))
ctmc = path2ctmc(xy=path$xy, t=path$t, rast=grad.stack)

```

Convert data so that we can use a glm to estimate resource parameters.

```{r}
# Brilliant! Hanks encodes the CRW term for you which accounts for the autocorrelation in the random walk.
glm_data = ctmc2glm(ctmc, loc.stack, grad.stack)

# Extract the hour of the day to look for time varying covariates
glm_data$hours = hour(as.POSIXct(glm_data$t, origin = '1970-01-01', tz = 'GMT'))
glm_data$forest_loc_hours = glm_data$forest_loc * glm_data$hours
glm_data$crw_hours = glm_data$crw * glm_data$hours

```


Fit various GLMs for both time-varying and time-invariant resource effects.

```{r}
# Don't include the gradient information as it is not useful as zeros and 1s
library(glmnet)
df = 8

fit = glm(z ~ forest_loc + crw + elev_grad,
              family="poisson", offset=log(tau), data=glm_data)

# Build time-dependent design matrix
bs_hours = bs(glm_data$hours, df=df) # B-spline basis for time effect
bs_hours_forest_loc = glm_data$forest_loc * bs_hours
bs_hours_crw = glm_data$crw * bs_hours
bs_hours_elev_grad = glm_data$elev_grad * bs_hours
X = cbind(1, bs_hours, bs_hours_forest_loc, bs_hours_crw, bs_hours_elev_grad)
colnames(X) = c("Intercept", paste0("hour_", 1:df), paste0("hforest_", 1:df), paste0("hcrw_",  1:df),
                 paste0("helevgrad_",  1:df))

# fit_time1 and fit_time2 are the same, just showing how to build the basis design matrix
fit_time1 = glm(z ~ bs(hours, df=df) + forest_loc:bs(hours, df=df) + elev_grad:bs(hours, df=df) + crw:bs(hours, df=df),
              family="poisson", offset=log(tau), data=glm_data)

fit_time2 = glm.fit(x=X, y=glm_data$z, family=poisson(link="log"), offset=log(glm_data$tau))

fit_timecv = cv.glmnet(X, y=glm_data$z, family="poisson", offset=log(glm_data$tau), alpha=1)
fit_cv = cv.glmnet(model.matrix(fit), y=glm_data$z, family="poisson", offset=log(glm_data$tau), alpha=1)


#fit_time2 = gam(z ~ s(hours) + forest_loc:s(hours) + crw:s(hours), family="poisson", offset = log(tau), data=glm_data)

```


```{r}
summary(fit)
```

The effect of elevation gradient should be interpreted as the effect of the negative gradient.  In other words, pigs are more likely to go in the direction of the negative gradient of elevation (i.e. downhill).

```{r}
plot(fit_timecv) # Cross-validation plot. Looking for lambda that minimizes deviance
plot(fit_timecv$glmnet.fit, xvar='lambda') # Looking at the shrinkage of the coefficients
best = which(fit_timecv$glmnet.fit$lambda == fit_timecv$lambda.min) # Find which lambda minimizes
cv_betas = fit_timecv$glmnet.fit$beta[, 45] # Extract those coefficients
cv_betas
```

```{r}
library(car)
Anova(fit_time1)
```

```{r}
# Plot the basis effects
mod_coefs = cv_betas #fit_time1$coefficients
hour_betas = mod_coefs[2:9]
forest_betas = mod_coefs[10:17]
crw_betas = mod_coefs[18:25]
elev_betas = mod_coefs[26:33]
hrvals = bs_hours %*% t(t(hour_betas)) + mod_coefs[1]
frvals = bs_hours %*% t(t(forest_betas))
crwvals = bs_hours %*% t(t(crw_betas))
elevvals = bs_hours %*% t(t(elev_betas))

od = order(glm_data$hours)
plot(glm_data$hours[od], hrvals[od], type="l", xlab="Hours", ylab="Hour effect")

jpeg("../results/timeforest.jpg",  width=400, height=300)
plot(glm_data$hours[od], frvals[od], type="l", xlab="Hours", ylab="Forest effect")
dev.off()
ggplot(data=NULL) + geom_line(aes(x=glm_data$hours[od], y=frvals[od])) + theme_bw() + xlab("Hour of day") + ylab("Forest effect")
ggsave("../results/timeforest.jpg", width=5, height=3)
plot(glm_data$hours[od], crwvals[od], type="l", xlab="Hours", ylab="CRW effect")
plot(glm_data$hours[od], elevvals[od], type="l", xlab="Hours", ylab="Elev. effect")
```

Evidence that pig moves slower through the forest habitat.  Consistent with
what other models predict.

Looking at the time-dependent parameters.  This shows that during the day, When pigs are in forest they tend to stay in forest relative
to other habitats.  However, in ths evening or the morning, it doesn't really matter.  Consistent with pigs using forest and cover and bedding
down.

## Step 4: Get the rate matrix 

```{r}

# loc.stack = stack(developed, forest, crop)
# grad.stack = stack(developed, forest, crop)
# names(loc.stack) = paste0(nms[-4], "_loc")
# names(grad.stack) = paste0(nms[-4], "_grad")

# Note that this CTMC is not positive recurrent as all forest cells have the same probability and the animal can wonder freely between those
# cells.  Pretty sure at least.  This would make sense why the stationary distribution looks so funky...there is none!
R = getR(fit, loc.stack, grad.stack)

```

Simulate pig movement on the landscape

```{r}
## Simulate movement of individual on the landscape
source("pigfxns.R")
set.seed(20)
start = (89 * 67)
temp_fit = fit
temp_fit$coefficients[4] = 100
traj = sim_traj(R, fit, start, steps=1000)
```

Look at a realization of the movement trajectory

```{r}
trajRast = forest
values(trajRast) = 0
values(trajRast)[traj$positions] = 200
path = rasterToPoints(trajRast, function(x) x == 200)

jpeg("../results/sim2.jpg")
plot(rcval1)
points(path[, "x"], path[, "y"], pch=18, cex=0.4)
dev.off()

```

```{r}
# Cumulative time elapsed
plot(cumsum(traj$time) / (60 * 60 * 24), 1:length(traj$time), type="l", xlab="Days", ylab="Events")
```

What does movement distance look like?

```{r}

# Unique cells visited
cells = extract(trajRast, path[, c('x', 'y')], cellnumbers=T)
locs = as.data.table(path[, c('x','y')])
locs$cells = cells[, 'cells']

obs = data.table(cells=traj$positions, time=c(cumsum(traj$time)))
obs = merge(obs, locs, key="cells")[order(time)]

```


```{r}
plot(obs$time, obs$x, type="l")
plot(obs$time, obs$y, type="l")
plot(obs$x, obs$y, type="l")
```


```{r}
longlat = obs[, c('x', 'y')]
names(longlat) = c('longitude', 'latitude')
dist = get_distance_vect(longlat)

plot(obs$time[-1] / (60 * 60 * 24), cumsum(dist), type="l", ylab="Cum. distance", xlab="Days")
```

Look at the utilization distribution

```{r}
temp_fit = fit
temp_fit$coefficients[4] = 20
R = getR(fit, loc.stack, grad.stack)
pi = get.UD(R)
ud = rcval1
values(ud) = pi

jpeg("../results/pi.jpg", width=300, height=300)
plot(log(ud))
dev.off()
```


```{r}
res = rast.grad(elcrop)
plot(res$rast.grad.x)
plot(elcrop)
```
This is the long-term utilization distribution based on elevation and forest


Now let's also look at the short term invasion probability based on the matrix exponential. R is the infintesimal generator matrix row-wise

```{r}
Q = R
diag(Q) = -1 * rowSums(R)

# Probability distribution at time t = 200
t = 200
p = rep(0, length(pi))
p[300] = 1

Pt = expm(Q*t)

```

## Running an analysis on all pigs in Texas within a certain time frame


```{r}
library(parallel)
source('pigfxns.R')

parallel_pigs = function(pigID, allpigs, numpoints, method, buffer, timestep){
  # Function for parallel processing pigIDs
  #
  # Parameters
  # ----------
  # pigID : vector of pigIDs
  # allpigs : data.table with all pig data
  # numpoints : the number of trajectory points for each pig
  
  print(paste("Working on pig", pigID))
  ind = (allpigs$pigID == pigID)
  pig = allpigs[ind][1:numpoints, ]
  colnames(pig)[c(8, 9, 13)] = c("y", "x", "time")
  
  # TODO: adjust file path
  rcval = raster("../data/raster_data/CDL_2016_clip_20180118163303_1712778437/CDL_2016_clip_20180118163303_1712778437.tif")
  el = raster("~/Downloads/n30w099/imgn30w099_1.img")
  
  # Project elevation raster 
  elproj = projectRaster(el, rcval)
  elproj = crop(elproj, extent(rcval))
  
  forest = dummy = rcval

  values(forest)[values(rcval) == 142] = 1 # Forest habitat
  values(forest)[values(rcval) != 142] = 0 # Not forest
  values(dummy) = 0 # Just use as a place holder
  
  loc.stack = stack(forest, dummy)
  names(loc.stack) = c("forest_loc", "dummy_loc") #paste0(nms, "_loc")
  grad.stack = stack(forest, dummy, elproj) #stack(developed, forest, crop, water)
  names(grad.stack) = c("forest_grad", "dummy_grad", "elev_grad") #paste0(nms, "_grad")
  
  glmdat = fit_ctmcmodel(pig, loc.stack, grad.stack, method=method, buffer=buffer, timestep=timestep)
  return(glmdat)
  
}
```

For all individuals, look at how the forest effect changes with time step. It should definitely decrease with increasing temporal scale...which it does.   At larger timescales, there shouldn't be much effect of forest. 

```{r}
timesteps = c("15 mins") #c("15 mins", "30 mins", "60 mins")
forest_coefs = list()
pigIDs = unique(pigdt$pigID)

for(i in 1:length(timesteps)){

  print(paste("Working on timestep:", timesteps[i]))
  allglm = mclapply(pigIDs, parallel_pigs, pigdt, 300, "interp", 0.001, timesteps[i], mc.cores = 4)
  glmdt = as.data.table(do.call(rbind, allglm))
  fitall_glm = glm(z ~ forest_loc*ID + crw + elev_grad, family=poisson, offset = log(as.numeric(tau)), data=glmdt) #tau is interevent time
  forest_coefs[[i]] = fitall_glm$coefficients
  
}
```

```{r}
forest_coefs
```

```{r}
vals = c(forest_coefs[[1]][2], forest_coefs[[1]][27:48] + forest_coefs[[1]][2])
ggplot(data=NULL) + geom_boxplot(aes(y=vals, x="All pigs")) + theme_bw() + ylab("Forest Effect") + xlab("n = 23 pigs collared in March") + 
    scale_y_continuous(breaks = round(seq(min(vals), max(vals), by = 0.5),1))
ggsave("../results/bplot.jpg", width=3, height=4)

```


```{r}
summary(fitall_glm)
```

Pretty robust to the time step that we are using.  Pigs move slower in forest and tend to stay there. Though, see below, this varies with time of day AND with pig.
Moreover, pigs tend to continue to move in the 

Let's look at the time-dependent effect of forest for the texas pigs.

```{r}
allglm = mclapply(pigIDs, parallel_pigs, testpigsdt, 300, "interp", 0.001, "15 mins", mc.cores = 4)

```

```{r}
glmdt = as.data.table(do.call(rbind, allglm))

glmdt$hours = hour(as.POSIXct(as.numeric(glmdt$t), origin = '1970-01-01', tz = 'GMT'))

IDs = unique(glmdt$ID)
res = list()
for(i in 1:length(IDs)){
  
  # Time-dependent forest effects between individuals
  df = 8
  tfit = glm(z ~ bs(hours, df=df) + forest_loc:bs(hours, df=df) + crw:bs(hours, df=df),
              family="poisson", offset=log(as.numeric(tau)), data=glmdt[glmdt$ID == IDs[i], ])
  
  bshours = bs(glmdt$hours, df=df)
  predbasis = predict(bshours, newx=0:23)
  forest_pred = predbasis %*% t(t(tfit$coefficients[10:17]))
  res[[i]] = data.frame(pred=forest_pred, ID=IDs[i], hour=0:23)
  
}

forest_pred = do.call(rbind, res)
ggplot(forest_pred) + geom_line(aes(x=hour, y=pred, color=ID)) + theme_bw() + ylab("Forest effect") + xlab("Hour of day")

```

There is variation between individuals regarding the the time-dependent effect of forest.  In general, forest has a negative effect, such that individuals tend to stay in the forest once they are there.  However, some individuals don't show this effect as strongly.


```{r}
library(lme4)
glmdt = as.data.table(do.call(rbind, allglm))
fitall = glmer(z ~ forest_loc + crw + (1 | ID), family=poisson, offset = log(tau), data=glmdt) # tau is interevent time
fitall_glm = glm(z ~ forest_loc + crw + ID, family=poisson, offset = log(tau), data=glmdt) #tau is interevent time
```

```{r}
summary(fitall)
summary(fitall_glm)
anova(fitall_glm, update(fitall_glm, .~. - ID), test="Chisq")
```


