---
title: "CTMC model of Tejon pigs"
author: Mark Wilber and Sarah Chinn
output:
  html_notebook: default
  pdf_document: default
bibliography: /Users/mqwilber/Dropbox/Documents/Bibfiles/Projects_and_Permits-feral_swine.bib
---

```{r, echo=FALSE, include=FALSE}
library(data.table)
library(raster)
library(ggplot2)
library(lubridate)
library(sp)
library(rgdal)
library(rgeos)
library(ctmcmove)
library(splines)
library(fda)
source('/Users/mqwilber/Repos/rsf_swine/code/pigfxns.R')
```

## Step 1: Identify outliers using K-means

```{r, echo=FALSE, include=FALSE}

## Clean the Camp Bullis Data

# For each pig, remove unrealistic observations
dat = fread("../data/formatted/full_pig_data.csv")
tcdat = dat[study == "tejon"]
rm(dat) # Free up some space

```

```{r}
# At least three pigs have some errant movements
tplot = ggplot(tcdat) + geom_path(aes(x=longitude, y=latitude)) + facet_wrap(~pigID)
tplot
```


```{r}
# There are a few pigs where we seem to have errant movements

badpigs = c("tejonF02_F12", "tejonM314")

badout = array(NA, dim=2)
fullind = list()

for(i in 1:length(badpigs)){
  
  bp = badpigs[i]
  bpdat = tcdat[pigID == bp, list(longitude, latitude)]
  kmeans.result = kmeans(bpdat, 1)
  centers = kmeans.result$centers[kmeans.result$cluster, ]
  distances <- sqrt(rowSums((bpdat - centers)^2))
  
  outliers <- order(distances, decreasing=T)[1:6]

  print(outliers) 
  badout[i] = outliers[1]
  fullind[[i]] = sapply(1:length(outliers), function(j) which((tcdat$pigID == bp) & (tcdat$longitude == bpdat$longitude[outliers[j]]) & (tcdat$latitude == bpdat$latitude[outliers[j]])))
  
  plot(bpdat[,list(longitude, latitude)], pch=19, col=kmeans.result$cluster, cex=1)
  points(kmeans.result$centers[, c("longitude", "latitude")], col=1:3, pch=15, cex=2)
  points(bpdat[outliers, c("longitude", "latitude")], pch="+", col=4, cex=3)
  
}

# Remove 6 outliers for "tejonF02_F12" and 1 outlier for "tejonM314"
rminds = c(fullind[[1]], fullind[[2]][1])

tcdat[c(fullind[[1]], fullind[[1]] + 1, fullind[[1]] - 1)][order(datetime)]
```

```{r}
# Remove the substantial outliers.

trimdat = tcdat[-rminds, ]
tp = ggplot(trimdat) + geom_path(aes(x=longitude, y=latitude)) + facet_wrap(~pigID)
tp
```

## Step 2: Identify runs of data that are sufficient length

```{r}
# Now, given that behavior can change over the course of the season, we likely want to make inference over consistent time intervals. Let's look at a plot of collaring times.

trimdat$datetime = as.POSIXct(trimdat$datetime, tz="GMT")
collarDur = trimdat[, list(mindate=min(datetime), maxdate=max(datetime)), by=pigID]
tp = ggplot(data=collarDur) + geom_linerange(aes(x=pigID, ymin=mindate, ymax=maxdate)) + 
          theme(axis.text.x = element_text(angle = 90, hjust = 1,  size=3)) + 
          scale_y_datetime(breaks = seq(min(collarDur$mindate), max(collarDur$maxdate), by="month"))

tp
```

```{r}
# Which pigs are between have at least `clength` consecutive fixes of `ctime` minutes or less?

lrdat = trimdat[, list(lr=runs(datetime, ctime=200, clength=500)), by=pigID]
longdat = merge(trimdat, lrdat, by="pigID")[lr == TRUE]

collarDur2 = longdat[, list(mindate=min(datetime), maxdate=max(datetime)), by=pigID]
tp = ggplot(data=collarDur2) + geom_linerange(aes(x=pigID, ymin=mindate, ymax=maxdate)) + 
          theme(axis.text.x = element_text(angle = 90, hjust = 1,  size=10)) + 
          scale_y_datetime(breaks = seq(min(collarDur2$mindate), max(collarDur2$maxdate), by="month"))
tp
```

What are the maximum gaps in collar trajectories for each pig?

```{r}
diffunits = function(x){
  dt = diff(x)
  units(dt) = "mins"
  return(dt)
}

trimdat[, list(diff=max(diffunits(datetime))), by=pigID]
```

We are going to have instances where we have multiple runs for a single pig that are split by long gaps in between fix times.  In these cases, we will want to break these runs into independent trajectories

```{r}
maxtime = 200  # minutes
minlength = 200 # steps

unqpigs = unique(trimdat$pigID)

newdat = list()
for(pignm in unqpigs){
  
  tdat = trimdat[pigID == pignm]
  tdat$run_number = 0 
  runinds = split_runs(tdat$datetime, ctime=maxtime, clength=minlength)
  
  for(i in 1:length(runinds)){
    inds = runinds[[i]]
    tdat$run_number[inds] = i
  }
  
  keepinds = do.call(c, runinds)
  rundat = tdat[keepinds]
  newdat[[pignm]] = rundat
  
}

# This is data split up by consecutive runs.
newdat = do.call(rbind, newdat)
```

## Step 3: Fit the continuous time movement model to each run, for each individual


```{r}

unqpigs = unique(newdat$pigID)

# Loop through and fit a continuous time movement model for each pig on each run
allpaths = list()
for(pignm in unqpigs){
  
  tdat = newdat[pigID == pignm]
  predpaths = lapply(unique(tdat$run_number), function(x) continuous_path(tdat[run_number == x, list(x=longitude, y=latitude, datetime=datetime)], "15 mins", "interp" ))
  allpaths[[pignm]] = predpaths
  
}

```

```{r}

locvars = c("temperature", "croplayer", "ndvi", "landcover", "precipitation")
gradvars = c("ndvi", "landcover")
gradxyvar = c("croplayer")

landcovertypes = c("wetland", "canopycover")
distgrad_types = c("root_and_tuber", "fruit_and_nuts")

buffer = 0.001

for(pignm in unqpigs[1]){
  
  tdat = newdat[pigID == pignm]
  extobj = extent(c(xmin=min(tdat$longitude) - buffer, xmax=max(tdat$longitude) + buffer, ymin=min(tdat$latitude) - buffer, ymax=max(tdat$latitude) + buffer))
  
  loc_stack = process_covariates(locvars, "tejon", extobj, 
                                 min(tdat$datetime), max(tdat$datetime), ext="loc", 
                                 landcovertypes = landcovertypes)
  
  grad_stack = process_covariates(gradvars, "tejon", extobj,
                                 min(tdat$datetime), max(tdat$datetime), ext="grad",
                                 landcovertypes = landcovertypes)
  
  gradxy_stack = process_covariates(gradxyvar, "tejon", extobj,
                                 min(tdat$datetime), max(tdat$datetime), ext="grad",
                                 landcovertypes = landcovertypes, distgrad = TRUE, distgrad_types = distgrad_types)
  
  gradx_stack = lapply(gradxy_stack, function(m) m$xgrad)
  grady_stack = lapply(gradxy_stack, function(m) m$ygrad)
  
  # Ensure projections are the same...lose some edge values this way
  baseras = loc_stack[[paste('cereals_', year(min(tdat$datetime)), "_loc", sep="")]]
  loc_stackproj = stack(lapply(loc_stack, function(x) projectRaster(x, baseras,
                                                                                method="ngb")))
  grad_stackproj = stack(lapply(grad_stack, function(x) projectRaster(x, baseras,
                                                                                method="bilinear")))
  gradx_stackproj = stack(lapply(gradx_stack, function(x) projectRaster(x, baseras,
                                                                                method="ngb")))
  grady_stackproj = stack(lapply(grady_stack, function(x) projectRaster(x, baseras,
                                                                                method="ngb")))
  
  
  pigpaths = allpaths[[pignm]]
  
  # Loop through runs for individual pigs
  runglmdat = list()
  for(j in 1:length(pigpaths)){
    
    cat("Run", j, "\n")
    smallpath = pigpaths[[j]]
    
    # Get reduced extent
    minll = apply(smallpath$path[[1]], 2, min)
    maxll = apply(smallpath$path[[1]], 2, max)
    smallextobj = extent(c(xmin=minll[1] - buffer, xmax=maxll[1] + buffer, ymin=minll[2] - buffer, ymax=maxll[2] + buffer))
    
    tglmdat = fit_ctmcmodel(tdat, pignm, crop(loc_stackproj, smallextobj), crop(grad_stackproj, smallextobj), 
                            xygrad=TRUE, xgrad.stack=crop(gradx_stackproj, smallextobj), ygrad.stack = crop(grady_stackproj, smallextobj),
                            predPath=smallpath, path2ctmcMethod="LinearInterp")
    runglmdat[[j]] = tglmdat
     
  }
  pigglmdat = do.call(rbind, runglmdat)
  
}

```






```


