---
title: "Tejon analysis"
output: html_notebook
---

Trying validate model and covariates with a single pig and then scaling up to all pigs at Tejon Ranch.

```{r}
library(data.table)
library(raster)
library(yaml)
library(glmnet)
library(ggplot2)
```

Load in rasters and movement model

```{r}
anal_params = yaml.load_file("analysis_parameters.yml")
dat = fread("~/Repos/rsf_swine/results/glmdata_by_study/tejon.csv")
dat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
dat[ , hourofday:=scale(hour(datetime))]
dat[ , monthofyear:=month(datetime)]

# rawdat = fread("~/Repos/rsf_swine/data/formatted/full_pig_data.csv")
ras = raster("~/Repos/rsf_swine/data/covariate_data/croplayer/tejon/tejon_fruit_and_nuts_2015.tif")
cras = raster("~/Repos/rsf_swine/data/covariate_data/croplayer/tejon/tejon_cereals_2016.tif")
raswater = raster("~/Repos/rsf_swine/data/covariate_data/water/tejon/tejon_water_nndistance.tif")
rasel = raster("~/Repos/rsf_swine/data/covariate_data/elevation/tejon/tejon_elevation.tif")
rasndvi = raster("~/Repos/rsf_swine/data/covariate_data/ndvi/tejon/tejon_ndvi_4_2016.tif")
rasmast = raster("~/Repos/rsf_swine/data/covariate_data/masting/tejon/tejon_masting.tif")
rascover = raster("~/Repos/rsf_swine/data/covariate_data/landcover//tejon/tejon_canopycover.tif")
shpcrop = shapefile("~/Repos/rsf_swine/data/covariate_data/croplayer/tejon/tejon_fruit_and_nuts_2015.shp")
```

```{r}
cshp = shapefile("~/Repos/rsf_swine/data/covariate_data/water/tejon/raw/tejon_water_nwi_raw.shp")
cshp_trans = spTransform(cshp, CRS("+proj=longlat +datum=WGS84"))

# Only consider permanent and semi-permanent water sources: "H" and "F" in NWI lingo: https://www.fws.gov/wetlands/Data/Wetland-Codes.html
peren = cshp_trans[grepl("H", cshp_trans$ATTRIBUTE) | grepl("F", cshp_trans$ATTRIBUTE), ]
```

```{r}
unique(dat$pigID)
```

Visualize the pig movement 
```{r}
tdat = dat[pigID == "tejonM302"]
tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]

plot(crop(rascover, ras))
#plot(ras, col=c("white", "blue"), add=T)
plot(shpcrop, add=T)
plot(peren, add=T, border='blue')
points(tdat[, x.adj], tdat[, y.adj], cex=0.01, col="red", pch=19)
points(tdat[crop_loc == 1, x.adj], tdat[crop_loc == 1, y.adj], cex=0.1, col="gray", pch=19)


#points(tdat[, x.adj], tdat[, y.adj], cex=0.01, col="red", pch=19)
#points(tdat[crop_loc == 1, x.adj], tdat[crop_loc == 1, y.adj], cex=0.05, col="blue", pch=19)
#plot(sp, add=T)
```

## Fit a single, crop-using pig 

First just look at the effects of movement independent of hour

```{r}
require(doMC)
source("pigfxns.R")

tdat = dat[pigID == "tejonM302"]
tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]

Xfull = build_design_matrix2(tdat, c("ndvi_loc", "ndvi_grad", "masting_loc",
                                    "masting_grad", "water_grad",
                                    "elevation_loc", "elevation_grad",
                                    "fruit_and_nutsdists_grad",
                                    "canopycover_loc", "canopycover_grad", "crw"), 
                                    c("fruit_and_nuts_loc"))

registerDoMC(cores=4)

tfit = cv.glmnet(Xfull, y=tdat$z, offset=log(tdat$tau), 
                       family="poisson", 
                       alpha=1, nlambda=20, nfolds=4, parallel=TRUE)
```

```{r}
tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.min), drop=F]
```

Substantially slower movement once the pig is in a crop field. An overall tendency to move toward crop fields (NOTE: this is highly dependent on the time of day!).  A tendency to stay longer in higher cover 

Fit a stan model to this data

```{r}
library(rstan)

poisglm = stan_model("stan_files/poisson_glm.stan")
```

```{r}
Xfullstan = cbind(1, Xfull)
sdata = list(X=Xfullstan, N=nrow(Xfullstan), p=ncol(Xfullstan), z=tdat$z, tau=tdat$tau)
fitpglm = sampling(poisglm, data=sdata, chains=3, iter=2000)
```

Let's add in daily variation in movement behavior.

```{r}
tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
tdat[ , hourofday:=scale(hour(datetime))]

df_hour = 5
stdcols = c("ndvi_loc", "ndvi_grad", "masting_loc",
            "masting_grad", "water_grad",
            "elevation_loc", "elevation_grad",
            "fruit_and_nutsdists_grad",
            "canopycover_loc", "canopycover_grad", "crw")
nonstdcols = c("fruit_and_nuts_loc")
splinecols = c("ndvi_grad", "masting_grad", "elevation_grad", "fruit_and_nutsdists_grad",
                "canopycover_grad", "water_grad", "crw")
Xgam = build_daily_spline_design_matrix(tdat, stdcols, nonstdcols, splinecols, df_hour=df_hour)

registerDoMC(cores=4)
fitgam = cv.glmnet(Xgam, y=tdat$z, offset=log(tdat$tau), family="poisson", 
                       alpha=1, nlambda=20, nfolds=5, parallel=TRUE)
```

```{r}
bestbetas = fitgam$glmnet.fit$beta[, which(fitgam$lambda == fitgam$lambda.min), drop=T]
bestbetas
```

Plot the daily effects

```{r}
library(ggplot2)
splinehour = s(hourofday, bs="cc", k=df_hour)
pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday)))) # B-spline basis for hour effect

dailyres = list()
plist = c("base_hour_*", "ndvi_grad_*", "masting_grad_*", "elevation_grad_*", "fruit_and_nutsdists_grad_*",
                                                    "canopycover_grad_*", "water_grad_*", "crw_*")
#plist = c("canopycover_loc_hour_*")
for(pattern in plist) {
  
  eff = pred_hours %*% t(t(bestbetas[names(bestbetas) %like% pattern]))
  effdt = data.frame(hour=0:23, beta=eff, coef=pattern)
  
  dailyres[[pattern]] = effdt
}

dailyres = do.call(rbind, dailyres)
tplot = ggplot(dailyres) + geom_line(aes(x=hour, y=beta)) + xlab("Hour of day") + ylab("Effect of on pig movement") + theme_bw() + facet_wrap(~coef, scales="free")
print(tplot)
#ggsave("../docs/presentations/images/canopycoverloc.pdf", width=5, height =3)

```


This is exactly what we would expect based on a empirical exploration of the movement patterns.  A strong tendency to move toward the used fruit and nut fields during the evening hours and thi tendency reverses in the in the early morning hours where the pig is moving away from these fields. Because the fields are in low NDVI areas, there is a strong correlation 

---

Fit model to all pigs.

```{r}
unique(dat$pigID)
```

```{r}

allres = list()
nongam = list()
allbestbetas = list()

for(pig in unique(dat$pigID)){
  
  # Extract a format data
  tdat = dat[pigID == pig]
  cat("Fitting pig", pig, "\n")
  tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
  tdat[ , hourofday:=scale(hour(datetime))]
  
  # Set fngrad to 0 if it is NA
  if(all(is.na(tdat$fruit_and_nutsdists_grad)))
     fngrad = rep(0, nrow(tdat))
  else
    fngrad = tdat$fruit_and_nutsdists_grad
  
  tdat$fruit_and_nutsdists_grad = fngrad

  # Step 1: Fit the non-daily model
  Xfull = build_design_matrix2(tdat, c("ndvi_loc", "ndvi_grad", "masting_loc",
                                    "masting_grad", "water_grad",
                                    "elevation_loc", "elevation_grad",
                                    "fruit_and_nutsdists_grad",
                                    "canopycover_loc", "canopycover_grad", "crw"), 
                                    c("fruit_and_nuts_loc"))

  registerDoMC(cores=4)
  tfit = cv.glmnet(Xfull, y=tdat$z, offset=log(tdat$tau), 
                       family="poisson", 
                       alpha=1, nlambda=20, nfolds=4, parallel=TRUE)
  
  nongam[[pig]] = tfit$glmnet.fit$beta[, which(tfit$lambda == tfit$lambda.min), drop=T]
  
  # Step 2: Fit the daily model
  stdcols = c("ndvi_loc", "ndvi_grad", "masting_loc",
              "masting_grad", "water_grad",
              "elevation_loc", "elevation_grad",
              "fruit_and_nutsdists_grad",
              "canopycover_loc", "canopycover_grad", "crw")
  nonstdcols = c("fruit_and_nuts_loc")
  splinecols = c("ndvi_grad", "masting_grad", "elevation_grad", "fruit_and_nutsdists_grad",
                 "canopycover_grad", "water_grad", "crw")
  
  if(pig == "tejonM302")
    Xtemp = Xgam
  
  Xgam = build_daily_spline_design_matrix(tdat, stdcols, nonstdcols, splinecols, df_hour=5)
  
  registerDoMC(cores=4)
  fitgam = cv.glmnet(Xgam, y=tdat$z, offset=log(tdat$tau), family="poisson", 
                     alpha=1, nlambda=20, nfolds=5, parallel=TRUE)
  
  bestbetas = fitgam$glmnet.fit$beta[, which(fitgam$lambda == fitgam$lambda.1se), drop=T]
  allbestbetas[[pig]] = bestbetas
  
  # Step 3: Extract cyclic basis for hour effect
  splinehour = s(hourofday, bs="cc", k=df_hour)
  pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday))))
  
  dailyres = list()
  plist = c("base_hour_*", "ndvi_grad_*", "masting_grad_*", "elevation_grad_*", "fruit_and_nutsdists_grad_*",
                                                    "canopycover_grad_*", "water_grad_*", "crw_*")
  for(pattern in plist) {
    
    eff = pred_hours %*% t(t(bestbetas[names(bestbetas) %like% pattern]))
    effdt = data.frame(hour=0:23, beta=eff, coef=pattern)
    dailyres[[pattern]] = effdt
  }
  
  dailyres = do.call(rbind, dailyres)
  dailyres$pigID = pig
  cat("Saving", pig, "\n")
  allres[[pig]] = dailyres
  
}
```

```{r}
allres_dt = as.data.table(do.call(rbind, allres))
ggplot(allres_dt) + geom_line(aes(x=hour, y=beta, color=pigID)) + xlab("Hour of day") + ylab("Effect on hourly movement rate") + theme_bw() + facet_wrap(~coef, scales="free")
```

```{r}
locparams = sapply(allbestbetas, function(x) x[names(x) %like% "*_loc"])
mlocparams = melt(locparams)
colnames(mlocparams) = c("coef", "pig", "value")
ggplot(mlocparams) + geom_boxplot(aes(x=coef, y=value)) + theme_bw()
```

A consistent tendecy to remain longer in higher NDVI cells.  No consistent effect of elevation or masting. Considering how seasonal masting is, this is not altogether surprising. 
Canopy cover is, a bit more surprisingly much more variable. 

```{r}
nongamparams = do.call(rbind, nongam)
mparams = melt(nongamparams)
colnames(mparams) = c("pig", "coef", "value")

ggplot(mparams) + geom_boxplot(aes(x=coef, y=value)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

There are two pigs that seem to be doing some weird things: tejonF07 and tejonF05.  Let's look at these a bit more closely.

```{r}
require(lubridate)
tF07 = dat[pigID == "tejonF05"]
tF07[, datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]

max(tF07$datetime) - min(tF07$datetime)
plot(tF07$datetime, rep(1, nrow(tF07)))
hist(hour(tF07$datetime))
```

For these pigs, it does seem that they are moving much more during the day than other pigs in the population. Why is that? 

## Explore a PCA analysis on the variables for different pigs to look for simiarlities and differences


```{r}
library(ggfortify)
library(ggplot2)
allparams = do.call(rbind, allbestbetas)

# Drop columns that are all 0
not0 = !apply(allparams, 2, function(x) all(x == 0))
redparams = allparams[, not0]

redparams_dt = as.data.frame(redparams)
redparams_dt$pigID = rownames(redparams)

pcares = stats::prcomp(redparams, scale=T)
autoplot(pcares, data = redparams_dt, colour = 'pigID',
         loadings = T, loadings.colour = 'blue',
         loadings.label = T, loadings.label.size = 2, label=TRUE, label.size=3)


#(pcares$sdev)^2 / sum((pcares$sdev)^2)
#barplot(pcares$rotation[, 2][order(abs(pcares$rotation[, 2]))])
```

```{r}

locparams = redparams[ , colnames(redparams) %like% "*_loc"]

for(p in plist){
  sump = t(t(rowSums(redparams[, colnames(redparams) %like% p])))
  locparams = cbind(sump, locparams)
  colnames(locparams)[1] = p
}

locparams = locparams[, !(colnames(locparams) %in% c("crw_*", "base_hour_*"))]
locparams_dt = as.data.frame(locparams)
locparams_dt$pigID = rownames(locparams)

ploc = stats::prcomp(locparams, scale=T)
autoplot(ploc, data = locparams_dt, colour = 'pigID',
         loadings = T, loadings.colour = 'blue',
         loadings.label = T, loadings.label.size = 2, label=TRUE, label.size=3)

```

### Add in seasonal effects for Tejon pigs

We have considered daily effects of movement, no let's consider seasonal (i.e. monthly effects on movement).  Two approaches for doing this

1. Fit a montly spline
2. Break the data into 4 seasons and fit a "season" factor that interacts with daily effects.

**Predicted seasonal covariates**

Likely seasonal covariates

1. Crop gradient
2. NDVI (this is already varying monthly)
  - Gradient and location
3. Masting density
  - Gradient and location
4. Water usage
  - Gradient
  
```{r}
source("pigfxns.R")
tdat = dat[pigID == "tejonM302"]
tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
tdat[ , hourofday:=scale(hour(datetime))]
tdat[ , monthofyear:=month(datetime)]
seasons = list("summer" = c(6, 7, 8), "fall" = c(9, 10, 11), 
               "winter"= c(12, 1, 2), "spring"=c(3, 4, 5))

df_hour = 5
stdcols = c("ndvi_loc", "ndvi_grad", "masting_loc",
            "masting_grad", "water_grad",
            "elevation_loc", "elevation_grad",
            "fruit_and_nutsdists_grad",
            "canopycover_loc", "canopycover_grad", "crw")
nonstdcols = c("fruit_and_nuts_loc")
splinecols = c("ndvi_grad", "masting_grad", "elevation_grad", "fruit_and_nutsdists_grad",
                "canopycover_grad", "water_grad", "crw")
seasonalcols = c("fruit_and_nutsdists_grad", "masting_grad", "masting_loc", "ndvi_grad", "ndvi_loc", "water_grad")
source("pigfxns.R")
Xseason = build_seasonal_design_matrix(tdat, stdcols, nonstdcols, splinecols, seasonalcols, df_hour=df_hour)

registerDoMC(cores=4)
fitseason = cv.glmnet(Xseason, y=tdat$z, offset=log(tdat$tau), family="poisson", 
                       alpha=1, nlambda=20, nfolds=5, parallel=TRUE)

```


```{r}
bestbetas = fitseason$glmnet.fit$beta[, which(fitseason$lambda == fitseason$lambda.min), drop=T]
bestbetas
```

Plot the seasonal location effects

```{r}
# Which seasons was the pig in?
rsums = apply(Xseason[, names(seasons)], 2, sum)
inseason = names(rsums)[rsums != 0]

# Location effects
locparams = bestbetas[names(bestbetas) %like% "*loc*"]
unqparams = unique(do.call(c, lapply(strsplit(names(locparams), ":"), function(x) x[[1]])))

for(prm in unqparams){
  print(locparams[names(locparams) %like% prm])
}

```

```{r}
splinehour = s(hourofday, bs="cc", k=df_hour)
pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday)))) # B-spline basis for hour effect

dailyres = list()
plist = c("base_hour_*", "elevation_grad_*","canopycover_grad_*", "crw_*")
#plist = c("canopycover_loc_hour_*")
for(pattern in plist) {
  
  eff = pred_hours %*% t(t(bestbetas[names(bestbetas) %like% pattern]))
  effdt = data.frame(hour=0:23, beta=eff, coef=pattern)
  
  dailyres[[pattern]] = effdt
}

dailyres = do.call(rbind, dailyres)
tplot = ggplot(dailyres) + geom_line(aes(x=hour, y=beta)) + xlab("Hour of day") + ylab("Effect of on pig movement") + theme_bw() + facet_wrap(~coef, scales="free")
print(tplot)
```

```{r}
splinehour = s(hourofday, bs="cc", k=df_hour)
pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday)))) # B-spline basis for hour effect

dailyres = list()
rsums = apply(Xseason[, names(seasons)], 2, sum)
inseason = names(rsums)[rsums != 0]

plist = c("ndvi_grad_*", "masting_grad_*", "fruit_and_nutsdists_grad_*", "water_grad_*")
#plist = c("canopycover_loc_hour_*")
for(pattern in plist) {
  
  dailyres[[pattern]] = list()
  seasbetas = bestbetas[names(bestbetas) %like% pattern]
  
  for(season in inseason){
    eff = pred_hours %*% t(t(seasbetas[names(seasbetas) %like% season]))
    effdt = data.frame(hour=0:23, beta=eff, coef=pattern)
    effdt$season = season
    dailyres[[pattern]][[season]] = effdt
  }
}

seasonres = do.call(rbind, lapply(dailyres, function(x) do.call(rbind, x)))
tplot = ggplot(seasonres) + geom_line(aes(x=hour, y=beta, linetype=season)) + xlab("Hour of day") + ylab("Effect of on pig movement") + theme_bw() + facet_wrap(~coef, scales="free")
print(tplot)
```


## Fit seasonal model for all pigs

```{r}

source("pigfxns.R")

allseasbetas = list()
splinelist = list()
for(pig in unique(dat$pigID)){
  
  cat("Fitting", pig, "\n")
  tdat = dat[pigID == pig]
  tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
  tdat[ , hourofday:=scale(hour(datetime))]
  tdat[ , monthofyear:=month(datetime)]
  
  # Set fngrad to 0 if it is NA
  if(all(is.na(tdat$fruit_and_nutsdists_grad)))
     fngrad = rep(0, nrow(tdat))
  else
    fngrad = tdat$fruit_and_nutsdists_grad
  
  tdat$fruit_and_nutsdists_grad = fngrad
  
  
  seasons = list("summer" = c(6, 7, 8), "fall" = c(9, 10, 11), 
                 "winter"= c(12, 1, 2), "spring"=c(3, 4, 5))
  
  df_hour = 5
  stdcols = c("ndvi_loc", "ndvi_grad", "masting_loc",
              "masting_grad", "water_grad",
              "elevation_loc", "elevation_grad",
              "fruit_and_nutsdists_grad",
              "canopycover_loc", "canopycover_grad", "crw")
  nonstdcols = c("fruit_and_nuts_loc")
  splinecols = c("ndvi_grad", "masting_grad", "elevation_grad", "fruit_and_nutsdists_grad",
                 "canopycover_grad", "water_grad", "crw")
  seasonalcols = c("fruit_and_nutsdists_grad", "masting_grad", "masting_loc", "ndvi_grad", "ndvi_loc", "water_grad",
                    "crw")
  Xseason = build_seasonal_design_matrix(tdat, stdcols, nonstdcols, splinecols, seasonalcols, df_hour=df_hour)
  
  registerDoMC(cores=4)
  fitseason = cv.glmnet(Xseason, y=tdat$z, offset=log(tdat$tau), family="poisson", 
                        alpha=1, nlambda=20, nfolds=5, parallel=TRUE)
  
  bestbetas = fitseason$glmnet.fit$beta[, which(fitseason$lambda == fitseason$lambda.1se), drop=T]
  
  allseasbetas[[pig]] = bestbetas
  
  splinehour = s(hourofday, bs="cc", k=df_hour)
  pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday)))) # B-spline basis for hour effect
  
  dailyres = list()
  rsums = apply(Xseason[, names(seasons)], 2, sum)
  inseason = names(rsums)[rsums != 0]
  
  plist = c("ndvi_grad_*", "masting_grad_*", "fruit_and_nutsdists_grad_*", "water_grad_*", "crw_*", "base_hour_*")
  #plist = c("canopycover_loc_hour_*")
  for(pattern in plist) {
    
    dailyres[[pattern]] = list()
    seasbetas = bestbetas[names(bestbetas) %like% pattern]
    
    for(season in inseason){
      eff = pred_hours %*% t(t(seasbetas[names(seasbetas) %like% season]))
      effdt = data.frame(hour=0:23, beta=eff, coef=pattern)
      effdt$season = season
      dailyres[[pattern]][[season]] = effdt
    }
  }
  
  seasonres = do.call(rbind, lapply(dailyres, function(x) do.call(rbind, x)))
  seasonres$pigID = pig
  splinelist[[pig]] = seasonres
  
}

```

```{r, fig.width=12, fig.height=12}

allpigs_seasoncoef = as.data.table(do.call(rbind, splinelist))
allpigs_seasoncoef$season = factor(allpigs_seasoncoef$season, levels=c("winter", "spring", "summer", "fall"))
meanpigs = allpigs_seasoncoef[, list(meanbeta=mean(beta)), by=list(coef, season, hour)]
meanpigs$season = factor(meanpigs$season, levels=c("winter", "spring", "summer", "fall"))

tplot = ggplot(data=NULL) + geom_line(data=allpigs_seasoncoef, aes(x=hour, y=beta, color=pigID), alpha=0.5) + geom_line(data=meanpigs, aes(x=hour, y=meanbeta), size=1) + 
                    xlab("Hour of day") + ylab("Effect of on pig movement") + theme_bw() + facet_wrap(~coef + season, scales="free", nrow=6, ncol=4)
print(tplot)
```

Plot the variable location coefficients by season

```{r}
combetas = do.call(rbind, allseasbetas)

# Extract the location columns
combetas = combetas[ , colnames(combetas) %like% "*_loc*"]
mcombetas = melt(combetas)
colnames(mcombetas) = c("pig", "coef", "value")
mcombetas$season = sapply(strsplit(as.character(mcombetas$coef), ":"), function(x) x[2])
mcombetas$coef1 = sapply(strsplit(as.character(mcombetas$coef), ":"), function(x) x[1])

mcombetas$season[is.na(mcombetas$season)] = "nonseasonal"
mcombetas$season = factor(mcombetas$season, levels=c("winter", "spring", "summer", "fall", "nonseasonal"))
ggplot(mcombetas) + geom_boxplot(aes(x=season, y=value)) + facet_wrap(~coef1, scales="free") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Similar to the previous analysis.  We see an increased tendency to linger in areas with high masting tree density in the winter (though this might be driven by very few points). The remainder of the year there is not as much selection for areas with high masting density.


## Temperature and precipitation analysis

Instead of using a seasonal factor as we do in the previous analysis, let's replace this seasonal factor with an interaction between monthly temperature and precipitation and our seasonal variables.

```{r}
tdat = dat[pigID == "tejonM302"]
tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
tdat[ , hourofday:=scale(hour(datetime))]
tdat[ , monthofyear:=month(datetime)]

df_hour = 5
stdcols = c("ndvi_loc", "ndvi_grad", "masting_loc",
            "masting_grad", "water_grad",
            "elevation_loc", "elevation_grad",
            "fruit_and_nutsdists_grad",
            "canopycover_loc", "canopycover_grad", "crw")
nonstdcols = c("fruit_and_nuts_loc")
splinecols = c("ndvi_grad", "masting_grad", "elevation_grad", "fruit_and_nutsdists_grad",
               "canopycover_grad", "water_grad", "crw")
seasonalcols = c("fruit_and_nutsdists_grad", "masting_grad", "masting_loc", "ndvi_grad", "ndvi_loc", "water_grad",
                 "crw")
Xseason = build_tempprecip_design_matrix(tdat, stdcols, nonstdcols, splinecols, seasonalcols, df_hour=df_hour)

registerDoMC(cores=4)
fitseason = cv.glmnet(Xseason, y=tdat$z, offset=log(tdat$tau), family="poisson", 
                      alpha=1, nlambda=20, nfolds=5, parallel=TRUE)


```

```{r}
bestbetas = fitseason$glmnet.fit$beta[, which(fitseason$lambda == fitseason$lambda.1se), drop=T]
bestbetas
```

```{r, fig.width=10, fig.height=4}

# Build a monthly temperature and precipitation predictor
monthtemp = dat[, list(stemp=scale2(temperature_loc), monthofyear)][, list(temperature_loc=mean(stemp)), by=monthofyear]
monthprecip = dat[, list(sprecip=scale2(precipitation_loc), monthofyear)][, list(precipitation_loc=mean(sprecip)), by=monthofyear]
monthtp = cbind(monthtemp, monthprecip[, list(precipitation_loc)])
monthtp[, ("temperature_loc:precipitation_loc"):=temperature_loc*precipitation_loc]

splinehour = s(hourofday, bs="cc", k=df_hour)
pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday)))) # B-spline basis for hour effect

dailyres = list()
inmonth = unique(tdat$monthofyear)

plist = c("ndvi_grad_*", "masting_grad_*", "fruit_and_nutsdists_grad_*", "water_grad_*", "crw_*", "base_hour_*")
tpcols = c("temperature_loc", "precipitation_loc", "temperature_loc:precipitation_loc")
#plist = c("canopycover_loc_hour_*")

for(pattern in plist) { # Loop through each pattern
  
  dailyres[[pattern]] = list()
  tpbetas = bestbetas[names(bestbetas) %like% pattern]
  
  for(month in inmonth){
      
      basenames = paste0(strsplit(pattern, "*", fixed=T)[[1]], 1:(df_hour - 1))
      baseeff = pred_hours %*% t(t(tpbetas[basenames]))
      
      for(tp in tpcols){
        
        tnames = paste0(strsplit(pattern, "*", fixed=T)[[1]], 1:(df_hour - 1), ":", tp)
        tpval = as.numeric(monthtp[monthofyear == month, tp, with=F])
        teff = (pred_hours*tpval) %*% t(t(tpbetas[tnames]))
        
        baseeff = baseeff + teff
      }
        
      effdt = data.frame(hour=0:23, beta=baseeff, coef=pattern)
      effdt$monthofyear = month
      dailyres[[pattern]][[month]] = effdt
  }
}

monthres = do.call(rbind, lapply(dailyres, function(x) do.call(rbind, x)))
tplot = ggplot(monthres) + geom_line(aes(x=hour, y=beta, color=as.factor(monthofyear))) + xlab("Hour of day") + ylab("Effect on pig movement") + theme_bw() + facet_wrap(~coef, scales="free")

tpplot = ggplot(monthtp) + geom_line(aes(x=monthofyear, y=temperature_loc, color="temperature")) + 
                           geom_line(aes(x=monthofyear, y=precipitation_loc, color="precipitation")) + theme_bw() + xlab("Month of year") + ylab("Standardized mean monthy value")

require(gridExtra)
grid.arrange(tplot, tpplot, nrow=1, ncol=2)

```


Plot how the non-daily varying location-based variables change with temperature

```{r, fig.width=6, fig.height=2}
basenames = sapply(strsplit(names(bestbetas), ":", fixed=T), function(x) x[[1]][1])
locparams = bestbetas[basenames %like% "*_loc*"]

seas_lps = c("ndvi_loc", "masting_loc") 


loceffects = list()
for(param in seas_lps){
  
  loceffects[[param]] = list()
  
  for(month in inmonth){
    
    baseeff = locparams[param]
    
    for(tp in tpcols){
      
      tpval = as.numeric(monthtp[monthofyear == month, tp, with=F])
      baseeff = baseeff + tpval*locparams[paste0(param, ":", tp)]
      
    }
    
    loceffects[[param]][[month]] = data.frame(month=month, coef=param, beta=baseeff)
  }
  
}

monthloc = do.call(rbind, lapply(loceffects, function(x) do.call(rbind, x)))
tplot = ggplot(monthloc) + geom_path(aes(x=month, y=beta)) + xlab("Month of year") + ylab("Effect on pig movement") + theme_bw() + facet_wrap(~coef)
#tplot

tpplot = ggplot(monthtp) + geom_line(aes(x=monthofyear, y=temperature_loc, color="temperature")) + 
                           geom_line(aes(x=monthofyear, y=precipitation_loc, color="precipitation")) + theme_bw() + xlab("Month of year") + ylab("Standardized mean monthy value")

require(gridExtra)
grid.arrange(tplot, tpplot, nrow=1, ncol=2)

```

## Analysis on all pigs

```{r}

alltpres = list()
alltpbestbetas = list()

for(pig in unique(dat$pigID)){
  
  cat("Working on", pig, "\n")
  tdat = dat[pigID == pig]
  tdat[ , datetime:=as.POSIXct(t, origin = '1970-01-01', tz = 'GMT')]
  tdat[ , hourofday:=scale(hour(datetime))]
  tdat[ , monthofyear:=month(datetime)]
  
  # Set fngrad to 0 if it is NA
  if(all(is.na(tdat$fruit_and_nutsdists_grad)))
     fngrad = rep(0, nrow(tdat))
  else
    fngrad = tdat$fruit_and_nutsdists_grad
  
  tdat$fruit_and_nutsdists_grad = fngrad
  
  df_hour = 5
  stdcols = c("ndvi_loc", "ndvi_grad", "masting_loc",
              "masting_grad", "water_grad",
              "elevation_loc", "elevation_grad",
              "fruit_and_nutsdists_grad",
              "canopycover_loc", "canopycover_grad", "crw")
  nonstdcols = c("fruit_and_nuts_loc")
  splinecols = c("ndvi_grad", "masting_grad", "elevation_grad", "fruit_and_nutsdists_grad",
                 "canopycover_grad", "water_grad", "crw")
  seasonalcols = c("fruit_and_nutsdists_grad", "masting_grad", "masting_loc", "ndvi_grad", "ndvi_loc", "water_grad",
                   "crw")
  Xseason = build_tempprecip_design_matrix(tdat, stdcols, nonstdcols, splinecols, seasonalcols, df_hour=df_hour)
  
  registerDoMC(cores=4)
  fitseason = cv.glmnet(Xseason, y=tdat$z, offset=log(tdat$tau), family="poisson", 
                        alpha=1, nlambda=20, nfolds=5, parallel=TRUE)
  
  bestbetas = fitseason$glmnet.fit$beta[, which(fitseason$lambda == fitseason$lambda.1se), drop=T]
  
  # Build a monthly temperature and precipitation predictor
  monthtemp = dat[, list(stemp=scale2(temperature_loc), monthofyear)][, list(temperature_loc=mean(stemp)), by=monthofyear]
  monthprecip = dat[, list(sprecip=scale2(precipitation_loc), monthofyear)][, list(precipitation_loc=mean(sprecip)), by=monthofyear]
  monthtp = cbind(monthtemp, monthprecip[, list(precipitation_loc)])
  monthtp[, ("temperature_loc:precipitation_loc"):=temperature_loc*precipitation_loc]
  
  splinehour = s(hourofday, bs="cc", k=df_hour)
  pred_hours = Predict.matrix(smooth.construct2(splinehour, tdat, NULL), data.frame(hourofday=sort(unique(tdat$hourofday)))) # B-spline basis for hour effect
  
  dailyres = list()
  inmonth = unique(tdat$monthofyear)
  
  plist = c("ndvi_grad_*", "masting_grad_*", "fruit_and_nutsdists_grad_*", "water_grad_*", "crw_*", "base_hour_*")
  tpcols = c("temperature_loc", "precipitation_loc", "temperature_loc:precipitation_loc")
  #plist = c("canopycover_loc_hour_*")
  
  for(pattern in plist) { # Loop through each pattern
    
    dailyres[[pattern]] = list()
    tpbetas = bestbetas[names(bestbetas) %like% pattern]
    
    for(month in inmonth){
      
      basenames = paste0(strsplit(pattern, "*", fixed=T)[[1]], 1:(df_hour - 1))
      baseeff = pred_hours %*% t(t(tpbetas[basenames]))
      
      for(tp in tpcols){
        
        tnames = paste0(strsplit(pattern, "*", fixed=T)[[1]], 1:(df_hour - 1), ":", tp)
        tpval = as.numeric(monthtp[monthofyear == month, tp, with=F])
        teff = (pred_hours*tpval) %*% t(t(tpbetas[tnames]))
        
        baseeff = baseeff + teff
      }
      
      effdt = data.frame(hour=0:23, beta=baseeff, coef=pattern)
      effdt$monthofyear = month
      dailyres[[pattern]][[month]] = effdt
    }
  }
  
  monthres = do.call(rbind, lapply(dailyres, function(x) do.call(rbind, x)))
  monthres$pigID = pig
  alltpres[[pig]] = monthres
  alltpbestbetas[[pig]] = bestbetas
  
}
```

```{r, fig.width=8, fig.height=10}
allpigstp = as.data.table(do.call(rbind, alltpres))
meanpig = allpigstp[, list(beta=mean(beta)), by=list(monthofyear, hour, coef)]
meanpig$pigID = "meantejonpig"
meanpig = meanpig[, colnames(allpigstp), with=F]
allpigstpfull = rbind(allpigstp, meanpig)


cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
tplot = ggplot(allpigstpfull) + geom_line(aes(x=hour, y=beta, color=as.factor(monthofyear))) + xlab("Hour of day") + ylab("Effect on pig movement") + theme_bw() + facet_wrap(~pigID + coef, scales="free", ncol=6, nrow=14) #+ scale_colour_brewer(palette = "Spectral")
tplot

```




Plot the non-daily coefficients

```{r, fig.width=8, fig.height=4}

allpigloceffects = list()
for(pig in names(alltpbestbetas)){
  
  inmonth = unique(dat[pigID == pig]$monthofyear)
  bestbetas = alltpbestbetas[[pig]]
  basenames = sapply(strsplit(names(bestbetas), ":", fixed=T), function(x) x[[1]][1])
  locparams = bestbetas[basenames %like% "*_loc*"]
  
  seas_lps = c("ndvi_loc", "masting_loc") 
  
  loceffects = list()
  for(param in seas_lps){
    
    loceffects[[param]] = list()
    
    for(month in inmonth){
      
      baseeff = locparams[param]
      
      for(tp in tpcols){
        
        tpval = as.numeric(monthtp[monthofyear == month, tp, with=F])
        baseeff = baseeff + tpval*locparams[paste0(param, ":", tp)]
        
      }
      
      loceffects[[param]][[month]] = data.frame(month=month, coef=param, beta=baseeff)
    }
    
  }
  
  monthloc = do.call(rbind, lapply(loceffects, function(x) do.call(rbind, x)))
  monthloc$pigID = pig
  allpigloceffects[[pig]] = monthloc
}

loccoefs = as.data.table(do.call(rbind, allpigloceffects))
mloccoefs = loccoefs[, list(beta=median(beta)), by=list(coef, month) ]
tplot = ggplot(loccoefs) + geom_point(aes(x=month, y=beta)) + geom_smooth(aes(x=month, y=beta), stat="smooth") + xlab("Month of year") + ylab("Effect on pig movement") + theme_bw() + facet_wrap(~coef)
tplot

tpplot = ggplot(monthtp) + geom_line(aes(x=monthofyear, y=temperature_loc, color="temperature")) + 
                           geom_line(aes(x=monthofyear, y=precipitation_loc, color="precipitation")) + theme_bw() + xlab("Month of year") + ylab("Standardized mean monthy value")

require(gridExtra)
grid.arrange(tplot, tpplot, nrow=1, ncol=2)


```


Wow, really hard to see much consistency. 


